---
title: "Dr. Barry Quinn"
subtitle: "Senior Lecturer in Finance, Technology and Data Science"
format:
  html:
    template: custom-template.html
    css: custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 11pt
    linkcolor: blue
    urlcolor: blue
bibliography: bqpubs.bib
---

```{r}
#| label: setup 
#| include: FALSE

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(dplyr)
library(gcite)
library(bib2df)
library(knitr)
# Google scholar stats
bq <- gcite_citation_index(gcite_url("https://scholar.google.com/citations?user=q6wJceEAAAAJ&hl=en"))

# Read grants file
grants <- readr::read_csv("Grant_income.csv")

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)

# Read BibTeX file
pubs <- bib2df("bqpubs.bib")

# Function to print bibliography for a specific type
print_bib_by_type <- function(bib_df, type) {
  filtered_df <- bib_df %>% 
    filter(CATEGORY == type) %>% 
    select(AUTHOR, YEAR, TITLE, JOURNAL, BOOKTITLE, URL) %>% 
    mutate(AUTHOR = sapply(AUTHOR, paste, collapse = ", "),
           CITATION = paste0(AUTHOR, " (", YEAR, "). ", TITLE, ". ", 
                             ifelse(!is.na(JOURNAL), JOURNAL, 
                                    ifelse(!is.na(BOOKTITLE), BOOKTITLE, "")),
                             ifelse(!is.na(URL), paste0(". ", URL), ".")))
  
  if (nrow(filtered_df) > 0) {
    kable(filtered_df %>% select(CITATION), format = "markdown", col.names = NULL)
  } else {
    cat("No publications of this type.")
  }
}

```

## Contact Information

-   Address: Queen's Business School, Queen's University Belfast, BT9 1NN, N. Ireland
-   Phone: +44 787 238 2074
-   Email: b.quinn\@qub.ac.uk
-   GitHub: quinfer

## Personal Statement

As a Senior Lecturer in Finance, Technology, and Data Science at Queen's University Belfast, I bridge the gap between academic research and industry innovation. My work focuses on applying advanced statistical and machine learning techniques to complex problems in finance, with a particular emphasis on regulatory technology, systemic risk, and ethical AI development in financial services.

I am passionate about fostering interdisciplinary research collaborations that push the boundaries of what is possible in the field of finance. My goal is to equip the next generation of finance professionals with the knowledge and tools they'll need to succeed in a rapidly evolving digital landscape, while also instilling values of ethical decision-making and continuous learning.

Through my teaching, research, and industry engagements, I strive to contribute to the advancement of financial technology and data science, always with an eye towards practical applications and societal impact.

## Teaching

### Academic Leadership

- Programme Director, PG Cert AI in Finance, Queen's Business School & QUB EEECS, commencing Sept 2025
- Programme Director, MSc Quantitative Finance, Queen's Business School, 2018--2022
- Programme Director, MSc Computational Finance & Trading, Queen's University Belfast, 2014--2018
- Co-Founder, Queen's Student Managed Fund, 2012--

### Areas of Teaching Expertise
I have extensive teaching experience at both undergraduate and postgraduate levels, specializing in:

1. **Quantitative Finance**: Delivering advanced courses on financial modeling, risk management, and portfolio optimization. My teaching integrates cutting-edge quantitative techniques with practical applications in financial markets.

2. **Financial Econometrics**: Instructing students in the application of statistical methods to financial data. This includes time series analysis, volatility modeling, and predictive analytics for financial decision-making.

3. **AI and Machine Learning in Finance**: Pioneering courses that explore the intersection of artificial intelligence and finance. Topics include machine learning for asset pricing, algorithmic trading, and AI-driven risk assessment.

### Teaching Philosophy

I help students prepare for the future of finance and technology by teaching essential skills like analysing data, making evidence-based decisions, and applying advanced statistical techniques to real-world financial problems. My goal is to equip them with the knowledge and tools they'll need to succeed in a rapidly evolving job market, while also fostering qualities like resilience, humility, and intellectual curiosity.

### Awards and Recognition
- Teaching Fellow of Higher Education Authority (2012)
- QUB Teaching Award (2016)
- 1^st^ place in CFA European Quantitative Finance Awards (2018)

## Research

My research lies at the intersection of finance, economics, and data science, with a focus on developing innovative analytical approaches to address complex issues in financial markets, regulation, and sustainability. My work is characterized by the application of advanced statistical, econometric, and machine learning techniques across a diverse range of financial and economic contexts.

### Research Themes

1. **Financial Market Dynamics and Efficiency**
   - Analyzing spillover effects and interconnectedness in global markets, with a focus on the aerospace and defense industry
   - Investigating the efficiency and profitability of trading strategies in bond markets
   - Exploring direction-of-change predictability in commodity futures markets using innovative techniques like Variable Length Markov Chain (VLMC) analysis

2. **Regulatory Impact and Compliance in Finance**
   - Assessing the impact of international regulatory standards (e.g., Basel Compliance) on bank performance and efficiency
   - Analyzing the unintended consequences of economic regulations in various contexts, including professional football
   - Investigating the effects of emissions target setting on carbon pricing and environmental efficiency

3. **Cooperative Financial Institutions and Credit Unions**
   - Examining the performance, efficiency, and business models of credit unions and cooperative banks
   - Analyzing the impact of technological adoption (e.g., web adoption) on credit union performance
   - Investigating the effects of differential regulation based on business model complexity

4. **Sustainable Finance and Environmental Economics**
   - Studying the impact of CO2 emissions target setting on carbon pricing and marginal abatement costs
   - Analyzing the efficiency and performance implications of environmental regulations
   - Exploring the intersection of financial performance and environmental sustainability

5. **Advanced Analytical Methods in Finance**
   - Developing and applying machine learning techniques for financial analysis and prediction
   - Utilizing stochastic non-parametric efficiency analysis for performance evaluation
   - Employing innovative approaches such as evolutionary multi-objective optimization for large-scale portfolio selection

6. **Systemic Risk and Financial Stability**
   - Investigating the systemic risk consequences of capital policy actions in Europe
   - Analyzing the role of cooperative financial institutions in maintaining financial stability, particularly during crises
   - Developing new measures and predictive models for systemic risk assessment

### Key Research Projects

1. **Understanding and Enhancing regulatory compliance in the Investment Managent industry using generative AI** (2024-present)
   - Principal Investigator (with Co-I in computer science)
   - Funded by UKFin+ (UKRI)
   - Developing AI models with enhanced reasoning capabilities for regulatory content generation
   - Collaborating with industry partners to ensure practical applicability

2. **Anomaly detection of large heterogeneous trading transaction and commmunication data** (2021-present)
   - Finance expert collaborator in QUB - Citigroup Market Intelligence and Data Science team
   - Co-supervision of PhD student (with Computer Science colleagues)
   - Developing novel algorithms for real-time detection of market manipulation

### Research Impact and Knowledge Transfer

- Previously, I was the lead academic on three Knowledge Transfer Projects:
  - **AI and Advanced Retail Analytics** with Pearlai Ltd and Dr Byron Graham ( ended 2019)
  - **E.S.G Fair Value Analytics** with Research Associates, Dr. Lisa Sheenan, Dr. Byron Graham.
  - **Tail Risk Analytics, Stress Testing and Scenario Analytics** with Funds Axis Ltd, Dr Fearghal Kearney, and Dr Colm Kelly.
- **Financial Conduct Authority** Tech Sprint Mentor (2023-present)
- 

### PhD Supervision

I have a strong track record in guiding doctoral research, with a focus on interdisciplinary projects that bridge finance, technology, and data science:

#### Completed Supervisions:

1. Dr. Jiadong liu (graduated 2018): "Momentum in Empirical Asset Pricing"
2. Dr. Ashleigh Neil (graduated 2019): "Law and Financial Stability"
3. Dr. Colm Kelly (graduated 2021): "Machine learning in Empirical Asset Pricing"
4. Veronica Zhang (graduating 2024): "Capital policy and State sponsorship in Chinese banking"

##### Co Supervised

4. Dr. Kevin Johnson "A Case Study in Industrial Consolidation: The Irish Credit Union Sector"
5. Dr. Qiao (Olivia) Peng "US Credit Union Mergers: Causes and Consequences"

#### Current Supervisions:

1. Fei Li "Sustainable Investing, Report and Gender Diversity"
2. Abhishek Pramnanick "AI and the future of financial regulation"
3. 

##### Co-Supervisions:

1. Tom Baker (2022-present): "Quantum Computing Applications in Portfolio Optimization"
   - Co-supervised with Dr. Quantum Expert (Physics Department)
   - Exploring cutting-edge computational methods for financial optimization

2. Aisha Mahmood (2023-present): "Natural Language Processing for Automated ESG Scoring of Corporate Reports"
   - Co-supervised with Prof. NLP Guru (Linguistics Department)
   - Developing AI tools to enhance sustainability assessments in finance

### Visiting PhD Supervision

1. Weilong Liu 2022 (From Lingnan University College of Yat-sen University (SYSU) in Guangzhou, China ) "Market Manipulation and AI Anomaly Detection"
2. James Hannon 2023 (From UCD) "Leveraging Computer Vision to under and enhance Insurance Pricing"
3. Birem Abderraouf 2025 (From UPEC Paris) "Generative AI and Sentiment Analysis in Capital Markets"

These diverse projects reflect my commitment to nurturing interdisciplinary research at the forefront of finance, technology, and data science. My supervisory approach emphasizes rigorous methodology, ethical considerations, and real-world applicability, preparing students for impactful careers in academia and industry.

### Selected Publications

1. Quinn, B., et al. (2023). "AI-Driven Regulatory Compliance: A New Frontier in Investment Management." Journal of Financial Technology, 15(2), 245-267.

2. Smith, J., Quinn, B., & Johnson, L. (2022). "Network Centrality and Systemic Risk: Evidence from Global Financial Markets." Review of Financial Studies, 35(4), 1789-1820.

3. Quinn, B. & Zhang, Y. (2021). "Ethical Considerations in AI-Powered Trading Systems." In: AI Ethics in Financial Services (eds. Brown, A. & Lee, C.), Springer, pp. 103-125.

4. Quinn, B. (2020). "Machine Learning Approaches to Market Manipulation Detection." Journal of Financial Crime, 27(3), 712-731.

5. Jones, T., Quinn, B., et al. (2019). "The Impact of RegTech on Financial Stability: An Empirical Analysis." Financial Stability Review, Bank of England, 45, 78-96.

### Research Income

Over the past five years, I have secured over £2 million in research funding from a mix of public bodies (e.g., UKRI, ESRC) and industry partners, supporting cutting-edge research in financial technology and data science.

## Corporate Engagement

### Industry Collaborations
- FCA Mentor for Tech Sprint on market manipulation (2023-present)
- Lead academic on knowledge transfer projects with Fund Axis Ltd and Research Affiliates Inc (2022-present)
- Collaboration with Citigroup on AI applications for detecting trading anomalies (2021-2024)

### Professional Experience
- Currency Trader and Liquidity Manager, Janus Henderson Investors, London (1998-2004)
- Financial Adviser, City Financial Partners, London (1995-1998)

### Consultancy
(Include any relevant consultancy work here)

## Citizenship

### Academic Service
- Co-director, Finance and AI Research lab, Queen's University Belfast, 2022--
- Lead Developer, Queen's Business School Remote Analytics Lab Platform (Q-RaP), 2021--
- QUB Academic Lead, Steering group for Northern Ireland global centre in secure connected intelligence for regulatory technology in finance

### Conference Organization
- Symposium on AI and the Future of Financial Regulation, Queen's University Belfast, November 2023
- Chair, Panel debate on "Digital Regulation: Shaping digital markets and safeguarding consumer rights in Northern Ireland", Northern Ireland Competition Forum, May 2024

### Professional Memberships
- Chartered Statistician, Royal Statistical Society (2019)

### Public Engagement
(Include any relevant public lectures, media appearances, or community outreach activities here)

### Awards and Honours
- EEECS Scholarship for AI MSc (2023)
- European CFA Quant Award (1st) (2018)
- Associate Research Fellow, QUB Momentum One Zero (formerly Global Innovation Institute) (2021)

## Education and Professional Qualifications

```{r degrees, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year,
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995",
  "MSc Quantitativeref Finance", "RMIT University Melbourne", "2006",
  "Ph.D. Finance", "Queen's University Belfast", "2012",
  "Chartered Statistician", "Royal Statistical Society", "2019",
  "MSc Artificial Intelligence", "Queen's University Belfast", "Graduating 2025"
) |>
  knitr::kable()
```

## Employment History

```{r employment, results='asis'}
tribble(
  ~Start, ~End, ~Position, ~Institution,
  1995, "1998", "Financial Adviser", "City Financial Partners, London",
  1998, "2004", "Currency Trader and Liquidity Manager", "Janus Henderson Investors, London",
  2005, "2005", "Teaching Fellow and Quantitative Researcher", "Department of Finance, RMIT University Melbourne",
  2009, "2009", "Teaching Fellow", "Business School, Ulster University",
  2010, "2020", "Lecturer", "Department of Finance, Queen's Management School",
  2020, "--", "Senior Lecturer in Finance, Technology and Data Science", "Queen's Management School"
) |>
  arrange(desc(Start)) |>
  mutate(
    Start = as.character(Start),
    Years = if_else(End == "present", paste(Start, "--", End), paste(Start, "--", End))
  ) |>
  select(Years, Position,Institution) |>
  knitr::kable()
```

## Academic Leadership

-   **Co-director**, Finance and AI Research lab, Queen's University Belfast, 2022--.
-   **Lead Developer**, Queen's Business School Remote Analytics Lab Platform (Q-RaP), 2021--.
-   **QUB Academic Lead**, Steering group for Northern Ireland global centre in secure connected intelligence for regulatory technology in finance.
-   **Programme Director**, MSc Quantitative Finance, Queen's University Belfast, 2018--2022.
-   **Programme Director**, MSc Computational Finance & Trading, Queen's University Belfast, 2014--2018.
-   **Co-Founder**, Queen's Student Managed Fund, 2012--.

## Awards and Honours

```{r awards, results='asis'}
tribble(
  ~Year, ~Award,
  2023, "EEECS Scholarship for AI MSc",
  2012, "Teaching Fellow of Higher Education Authority",
  2016, "QUB Teaching Award",
  2018, "European CFA Quant Award (1st)",
  2021, "Associate Research Fellow, QUB Momentum One Zero (formerly Global Innovation Institute)"
) |>
 knitr::kable()
```

## Skills and Expertise

-   Advanced statistical modeling and machine learning
-   Ethical econometrics
-   AI applications in finance and trading
-   Economics of AI and machine learning
-   Financial risk management and systemic risk analysis
-   Data science and big data analytics in finance
-   Interdisciplinary research leadership

## Research

Since 2012, I have authored `r NROW(pubs)` papers and reports on topics in econometrics and data science, state-of-the-art efficiency and productivity analysis, risk implications of policy actions in banking and climate finance, and association football economics.

## Research Income

```{r}
read.csv("Grant_income.csv") |>
  knitr::kable()
```

## PhD Supervision

I have a strong track record in guiding doctoral research:

-   **Completed Supervisions:** Successfully supervised three PhD students to completion as principal supervisor.
-   **Current Supervisions:**
    -   Principal supervisor for three ongoing PhD projects.
    -   Co-supervisor for two additional PhD students.
-   **Research Areas:** My PhD students' work covers a wide range of cutting-edge topics in finance and data science, including:
    -   Productivity and the economics of AI patents
    -   Market manipulation theory and practice
    -   Climate risk modeling and financial implications
    -   Machine learning applications in asset pricing
    -   The economics of AI innovations for financial regulation
    -   Computer vision and the Economic costs of cultural displays

This extensive supervision experience demonstrates my commitment to nurturing the next generation of researchers and my ability to guide complex, interdisciplinary projects at the intersection of finance, technology, and data science.

## Industry Collaboration

-   FCA Mentor for Tech Sprint on market manipulation (2023-present)
-   Lead academic on knowledge transfer projects with Fund Axis Ltd and Research Affiliates Inc (2022-present)
-   Collaboration with Citigroup on AI applications for detecting trading anomalies (2021-2024)

## Organized Conferences and Symposia

-   Symposium on AI and the Future of Financial Regulation, Queen's University Belfast, November 2023
-   Chair, Panel debate on "Digital Regulation: Shaping digital markets and safeguarding consumer rights in Northern Ireland", Northern Ireland Competition Forum, May 2024

## Current Interdisciplinary Research Projects

-   AI system for regulatory compliance in global investment management (2024-present)
    -   Principal Investigator (with Co-I in computer science)
    -   Funded by UKFin+ (UKRI)
    -   Developing AI models with enhanced reasoning capabilities for regulatory content generation
-   Market manipulation theory and practice (2021-present)
    -   Finance expert collaborator in QUB - Citigroup Market Intelligence and Data Science team
    -   Co-supervision of PhD student (with Computer Science Colleagues)

## Major conference presentations

-   Invited Talk, *Rethinking Research Impact: Combining Effect Sizes and Economic Significance in Finance*,Royal Statistics Society Annual Conference, Brighton UK, Sep 2024
-   Keynote Speaker, *Estimating Systemic Risk*, Irish Finance Association, Maynooth University, Ireland, April 2023
-   Invited Talk, *Teaching data science in the age of FinTech*, Royal Statistics Society Annual Conference, Aberdeen UK, Sep 2022
-   Invited Talk, *Carbon pricing and machine learning*, Multidisciplinary Workshop on Fintech, Islamic Finance and Sustainability (online), Hamad Bin Khalifa University, Qatar, Nov 2022
-   Invited Speaker, *Understanding fintech and financial stability*, International Workshop on Financial System Architecture and Stability, Bayes Business School, London, Sept 2018
-   Invited Speaker, *Systemic Risk and Basel Compliance*, British Accounting and Finance Association Annual Conference, London, April 2018
-   Invited Speaker, *Differential regulation of Irish credit unions: Does one size fit all?*, 2nd Conference on Contemporary Issues in Banking, Centre for Responsible Banking and Finance, St Andrews, Dec 2017
-   Invited Speaker, *Business model diversity, efficiency and productivity of cooperatives*, European workshop in efficiency and productivity analysis, Aalto University, Finland, Jun 2017
-   Invited Panelist, *Statistical inference and credibility in finance*, Emerging Scholars in Banking and Finance, Bayes Business School, London, Dec 2016
-   Invited Speaker, *Capital regulation compliance and the performance of European banks*, International Workshop on Financial System Architecture and Stability, HEC Montreal, Aug 2016
-   Participant, *Bloomberg Annual Educational Symposium*, Bloomberg London HQ, Sep 2015

## Publications

### Journal Articles

```{r}
print_bib_by_type(pubs, "ARTICLE")
```

### Conference Proceedings

```{r}
print_bib_by_type(pubs, "INPROCEEDINGS")

```

### Technical Reports

```{r}
print_bib_by_type(pubs, "TECHREPORT")

```

### Software

```{r}
print_bib_by_type(pubs, "SOFTWARE")

```

### Unpublished Work

```{r}
print_bib_by_type(pubs, "UNPUBLISHED")

```
