---
title: "Dr. Barry Quinn"
subtitle: "Senior Lecturer in Finance, Technology and Data Science"
format:
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 10pt
    linkcolor: blue
    urlcolor: blue
bibliography: bqpubs.bib
---

```{r}
#| label: setup 
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

-   Address: Queen's Business School, Queen's University Belfast, BT9 1NN, N. Ireland
-   Phone: +44 787 238 2074
-   Email: b.quinn\@qub.ac.uk
-   GitHub: quinfer

# Personal Statement

As a Senior Lecturer at Queen’s University Belfast, I have developed a robust academic and industry-oriented profile in financial technology. My work focuses on regulatory technology, systemic risk, and ethical AI applications in finance, aligning with the goals of centres such as Ulster University’s Centre of Financial and Regulatory Technology. I aim to enhance research impact, teaching excellence, and interdisciplinary collaboration, advancing institutional profiles in REF and beyond.

I have successfully supervised six PhD students to completion, with ongoing supervision of five more, reflecting my commitment to developing the next generation of researchers in financial technology and regulatory compliance.

## Education and Professional Qualifications

```{r degrees, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year,
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006",
  "Ph.D. Finance", "Queen's University Belfast", "2012",
  "Chartered Statistician", "Royal Statistical Society", "2019",
  "MSc Artificial Intelligence", "Queen's University Belfast", "Graduating 2025"
) |>
  knitr::kable()
```

# Teaching

I have extensive teaching experience at both undergraduate and postgraduate levels, specialising in:

1.  **Quantitative Finance**: My teaching integrates cutting-edge quantitative techniques with practical applications in financial markets.

2.  **Ethical Econometrics**: Instructing students in econometrics as an applied statistics discipline, avoiding the statistical cavaliering that plagues the econometric publication playbook.

3.  **AI and Machine Learning in Finance**: Pioneering courses that explore the intersection of artificial intelligence and finance. Topics include machine learning for rule-based algorithmic trading, asset pricing, algorithmic trading, and AI-driven risk assessment.

## Teaching Philosophy

I help students prepare for the future of finance and technology by teaching essential skills like analysing data, making evidence-based decisions under uncertainty, and applying advanced statistical techniques to real-world financial problems. My goal is to equip them with the knowledge and tools they'll need to succeed in a rapidly evolving job market, while also fostering qualities like resilience, humility, and intellectual curiosity.

# Research

I have published thirteen peer-reviewed internationally acclaimed academic papers (1 ABS4, 10 ABS3s, and 2 ABS2s), three commissioned research reports, and two statistical software research packages. My research lies at the intersection of finance, economics, and data science, with a focus on developing innovative analytical approaches to address complex issues in financial markets, regulation, and sustainability.

## Peer-reviewed Publications

-   [Liu, Weilong, Zhang, Yong, Liu, Kailong, Quinn, Barry, Yang, Xingyu, Peng, Qiao (2024). Evolutionary Multi-Objective Optimisation for Large-Scale Portfolio Selection With Both Random and Uncertain Returns. IEEE Transactions on Evolutionary Computation^ABS4^]{style="color: #2E8B57;"}

- <span style="color: #0066cc;">Bouri, E., Quinn, B., Sheenan, L. & Tang, Y. (2024) Investigating extreme linkage topology in the aerospace and defence industry. International Review of Financial Analysis^ABS3^</span>

-   [Quinn, Barry, Gallagher, Ronan, Kuosmanen, Timo (2023). Lurking in the shadows: The impact of CO2 emissions target setting on carbon pricing in the Kyoto agreement period Energy Economics^ABS3^]{style="color: #0066cc;"}

- <span style="color: #0066cc;">Liu, Jiadong, Papailias, Fotis, Quinn, Barry (2021). Direction-of-change forecasting in commodity futures markets. International Review of Financial Analysis^ABS3^</span>

-   [Gallagher, Ronan, Quinn, Barry (2020). Regulatory own goals: The unintended consequences of economic regulation in professional football. European Sport Management Quarterly^ABS3^]{style="color: #0066cc;"}

- <span style="color: #0066cc;">McKillop, Donal, French, Declan, Quinn, Barry, Sobiech, Anna L, Wilson, John OS (2020). Cooperative financial institutions: A review of the literature. International Review of Financial Analysis^ABS3^</span>

- <span style="color: #9370DB;">Quinn, Barry, Hanna, Alan, MacDonald, Fred (2018). Picking up the pennies in front of the bulldozer: The profitability of gilt based trading strategies. Finance Research Letters^ABS2^</span>

- <span style="color: #0066cc;">McKillop, Donal G, Quinn, Barry (2017). Irish credit unions: Differential regulation based on business model complexity. The British Accounting Review^ABS3^</span>

- <span style="color: #0066cc;">Ayadi, Rym, Naceur, Sami Ben, Casu, Barbara, Quinn, Barry (2016). Does Basel compliance matter for bank performance?. Journal of Financial Stability^ABS3^</span>


- <span style="color: #0066cc;">Glass, J, McKillop, Donal, Quinn, Barry (2015). Modelling the Performance of Irish Credit Unions, 2002-2010. Financial Accountability & Management^ABS3^</span>

- <span style="color: #9370DB;">McKillop, Donal G, Quinn, Barry (2015). Technology adoption by Irish credit unions: Performance implications. Annals of Public and Cooperative Economics^ABS2^</span>

- <span style="color: #9370DB;">Quinn, Barry, McKillop, Donal (2009). Internet banking and Irish credit unions. International Journal of Cooperative Management^ABS2^</span>

- <span style="color: #0066cc;">Glass, J Colin, McKillop, Donal G, Quinn, Barry, Wilson, John (2014). Cooperative bank efficiency in Japan: a parametric distance function analysis. The European Journal of Finance^ABS3^</span>

- <span style="color: #9370DB;">Quinn, Barry, McKillop, Donal (2009). Cost performance of Irish credit unions. Journal of Cooperative Studies^ABS2^</span>

## Software Publications


| Package | Year | Description | Repository |
|---------|------|-------------|------------|
| [Time Series Econometrics (tsfe)](https://doi.org/10.5281/zenodo.6376113) | 2022 | R Package for time series analysis | [GitHub](https://github.com/quinfer/tsfe) |
| [Financial Machine Learning (fml)](https://doi.org/10.5281/zenodo.6376114) | 2020 | R Package for financial ML | [GitHub](https://github.com/quinfer/fml) |


## Working Paper

-   Dai, Yongsheng, Wang, Hui, Rafferty, Karen, Spence, Ivor, Quinn, Barry. TDSRL: [*Time Series Dual Self-Supervised Representation Learning for Anomaly Detection from Different Perspectives*](https://ideas.repec.org/p/zbw/qmsrps/202403.html) (2nd R&R IEEE Internet of Things).

-   Liu, Kailong, McKillop, Donal, Quinn, Barry, Peng, Qiao. *Modelling and Predicting Credit Union Failures*. Centre for Responsible Banking and Finance Working Paper Series, St Andrew’s University (2nd R&R International Journal of Forecasting^ABS3^).

-   McKillop, Donal, Quinn, Barry, Peng, Qiao *Predicting Credit Union Merger Activity using Gaussian Process*

-   Quinn, Barry, Abhishek Pramanick *Regulatory Technology: Predicting Future Trends in the context of FinTech*

-   Quinn, Barry, Ronan Gallager *Great Expectations: Managerial Turnover and Market Expection in Association Football*

## Research Reports

-   Bryon Graham, Barry Quinn (2019). *Price comparison and web analytics*. (Commissioned by InvestNI)
-   Bryon Graham, Barry Quinn (2017). *Machine learning and predictive analytics in the mall retail business*. (Commissioned by InvestNI)
-   Fearghal Kearney, Barry Quinn (2020). The theoretical foundations of value at risk modelling. (Commissioned by InvestNI and Funds Axis Ltd)
-   French,Declan, McKillop, Donal, Quinn, Barry (2018). Landscape review of Northern Ireland credit unions.(Commissioned by Northern Ireland Department of Communities and Rural Affairs)

## Research Impact and Knowledge Transfer

1.  *Understanding and Enhancing Regulatory Compliance in the Investment Management Industry Using Generative AI* with Funds Axis Ltd and Momentum 1.0 (UKFin+ and UKRI, 2024-2025).
2.  *Anomaly Detection in Large Heterogeneous Trading Data* with Citigroup Belfast and Momentum 1.0 (2022-2023).
3.  *AI and Advanced Retail Analytics* with Pearlai Ltd and Dr. Byron Graham (KTP, 2017-2019).
4.  *E.S.G. Fair Value Analytics* with Research Associates, Dr. Lisa Sheenan, Dr. Byron Graham (KTP, 2021).
5.  *Tail Risk Analytics and Stress Testing* with Funds Axis Ltd, Dr. Fearghal Kearney, and Dr. Colm Kelly (KTP, 2021-2023).

## Current PhD Supervisions

To date, I have supervised six PhD students to completion, with ongoing supervision of six projects addressing cutting-edge topics in FinTech, ESG analytics, and generative AI. My current students include:

1. Fei Li: *Sustainable Investing, Report and Gender Diversity* (QBS Scholarship).
2. Abhishek Pramnanick: *AI and the Future of Financial Regulation* (QBS Scholarship).
3. Dongwei (Jeffery) Li: *The Economics of AI Patents and Productivity in Finance* (DofE Scholarship).
4. Yuqi Ding: *The Economics of Generative AI in Finance* (QBS Scholarship).
5. Yongshei Dai: *Market Manipulation and Financial Anomaly Detection* (PWC Scholarship).
6. Brandon Cochrane: *Exploring the Welfare Cost of Cultural Displays in Northern Ireland using multimodal Generative AI* (CAST Scholarship).

## Completed PhD Supervisions

1.  Dr. Jiadong liu (graduated 2018): *Momentum in Empirical Asset Pricing* **(Principal Supervisor)**
2.  Dr. Ashleigh Neil (graduated 2019): *Law and Financial Stability* **(Principal Supervisor)**
3.  Dr. Colm Kelly (graduated 2021): *Machine learning in Empirical Asset Pricing* **(Principal Supervisor)**
4.  Veronica Zhang (graduating 2024): *Capital policy and State sponsorship in Chinese banking* **(Principal Supervisor)**
5.  Dr. Kevin Johnson (graduated 2019) *A Case Study in Industrial Consolidation: The Irish Credit Union Sector* **(Co Supervised)**
6.  Dr. Qiao (Olivia) Peng (graduated 2021) *US Credit Union Mergers: Causes and Consequences* **(Co Supervised)**


## Visiting PhD Supervision

1.  Weilong Liu 2022 (From Lingnan University College of Yat-sen University (SYSU) in Guangzhou, China ) "Market Manipulation and AI Anomaly Detection"
2.  James Hannon 2023 (From UCD) "Leveraging Computer Vision to under and enhance Insurance Pricing"
3.  Birem Abderraouf 2025 (From UPEC Paris) "Generative AI in Capital Markets"

## Research Income

```{r}
read.csv("Grant_income.csv") %>%
    mutate(Amount = scales::comma(Amount, prefix = "£")) %>%
    knitr::kable(
      col.names = c("Source", "Role", "Start", "End", "People", "Title", "Full Economic Costs (£)", "Type"),
      align = c("l", "l", "l", "l", "l", "l", "r", "l"),
      format = "latex",
      caption = "Research Grant Income",
      booktabs = TRUE,
      escape = FALSE,
      longtable = TRUE  # Allows table to span multiple pages
    ) %>%
    kableExtra::kable_styling(
      font_size = 9,
      latex_options = c("repeat_header")
    ) %>%
    kableExtra::column_spec(1, width = "1.5cm") %>%  # Source
    kableExtra::column_spec(2, width = "1.5cm") %>%  # Role
    kableExtra::column_spec(3, width = "1.2cm") %>%  # Start
    kableExtra::column_spec(4, width = "1.2cm") %>%  # End
    kableExtra::column_spec(5, width = "2cm") %>%    # People
    kableExtra::column_spec(6, width = "4cm") %>%    # Title
    kableExtra::column_spec(7, width = "2cm") %>%    # Amount
    kableExtra::column_spec(8, width = "1.5cm")      # Type
```


# Professional Accreditations

- Chartered Statistician, Royal Statistical Society (2019).
- Advanced Data Science Professional, Alliance of Data Science Professionals (2024).

# Awards and Recognition

- Teaching Fellow of Higher Education Authority (2012).
- QUB Teaching Award (2016).
- 1st Place in CFA European Quantitative Finance Awards (2018).
- EEECS Scholarship for AI MSc (2023).

# Professional Experience

```{r employment, results='asis'}
tribble(
  ~Years, ~Position, ~Institution,
  "2020 - Present", "Senior Lecturer in Finance, Technology and Data Science", "Queen's University Belfast",
  "2010 - 2020", "Lecturer", "Queen's University Belfast",
  "2009", "Teaching Fellow", "Ulster University",
  "2005", "Teaching Fellow and Quantitative Researcher", "RMIT University, Melbourne",
  "1998 - 2004", "Currency Trader and Liquidity Manager", "Janus Henderson Investors, London",
  "1995 - 1998", "Financial Adviser", "City Financial Partners, London"
) |>
  knitr::kable()
```

# Skills and Expertise

- Advanced statistical modelling and machine learning.
- AI applications in finance and trading.
- Financial risk management and systemic risk analysis.
- Interdisciplinary research leadership.
