---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: ../assets/custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 11pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup 
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

- **Address**: Ulster University Business School, Belfast Campus, York Street, Belfast, BT1 5AB, N. Ireland
- **Email**: b.quinn1@ulster.ac.uk
- **GitHub**: [quinfer](https://github.com/quinfer)
- **ORCID**: [0000-0002-8637-9060](https://orcid.org/0000-0002-8637-9060)
- **Date of Birth**: 10 July 1973

# Personal Statement

Professor Barry Quinn is Professor of Finance & Financial Technology at Ulster University Business School, where he is developing the Centre for Financial and Regulatory Technology. He is an applied econometrician and Chartered Statistician whose work integrates econometrics, machine learning, and computational approaches to address challenges in financial markets, risk, and regulatory compliance.

His research combines causal policy analysis, algorithmic trading risk detection, and trustworthy AI methods, with a particular interest in how financial technology can support resilient markets and evidence-based regulation. Barry's work is collaborative and interdisciplinary, bridging economics, statistics, computer science, and law.

He teaches across quantitative finance, econometrics, and AI applications in finance, encouraging students to engage with rigorous, ethical, and reproducible methods. Pedagogically, he emphasises employability, resilience, and critical thinking.

Before entering academia, Barry worked in financial markets, specialising in currency trading and liquidity management. He holds a PhD in Finance from Queen's University Belfast, is a Chartered Statistician, and is completing an MSc in Artificial Intelligence. He actively collaborates with industry and policy partners to ensure his research delivers academic, societal, and regulatory impact.

## Education and Professional Qualifications

```{r degrees, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year,
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006",
  "Ph.D. Finance", "Queen's University Belfast", "2012",
  "Chartered Statistician", "Royal Statistical Society", "2019",
  "MSc Artificial Intelligence", "Queen's University Belfast", "2025"
) |>
  knitr::kable()
```

# Research Excellence

## Key Publications (Selected from 14 peer-reviewed papers)

**ABS4* Journal:**
- Liu, Weilong, Zhang, Yong, Liu, Kailong, Quinn, Barry, Yang, Xingyu, Peng, Qiao (2024). Evolutionary Multi-Objective Optimisation for Large-Scale Portfolio Selection With Both Random and Uncertain Returns. *IEEE Transactions on Evolutionary Computation* (ABS4)

**Recent ABS3 Publications:**
- McKillop, Donal, Liu, Kailong, Quinn, Barry, Peng, Qiao (Forthcoming). Modelling and Predicting Credit Union Failures. *International Journal of Forecasting* (ABS3)
- Bouri, E., Quinn, B., Sheenan, L. & Tang, Y. (2024). Investigating extreme linkage topology in the aerospace and defence industry. *International Review of Financial Analysis* (ABS3)
- Quinn, Barry, Gallagher, Ronan, Kuosmanen, Timo (2023). Lurking in the shadows: The impact of CO2 emissions target setting on carbon pricing in the Kyoto agreement period. *Energy Economics* (ABS3)

## Research Impact and Funding

```{r}
#| label: tbl-research-grants-academic
#| tbl-cap: "Major Research Grants (Selected)"

grants_academic <- tribble(
  ~Source, ~Role, ~Period, ~Title, ~Amount,
  "UKRI/InnovateUK", "PI", "2024-2025", "Understanding and Enhancing regulatory compliance using AI", "£250K",
  "InnovateUK", "PI", "2021-2023", "Tail Risk Analytics and Stress Testing", "£180K", 
  "InnovateUK", "Co-I", "2017-2019", "AI and Advanced Retail Analytics", "£150K",
  "Various KTPs", "PI/Co-I", "2017-2023", "Multiple Knowledge Transfer Partnerships", "£800K+"
)

grants_academic |>
  knitr::kable(
    col.names = c("Funding Body", "Role", "Period", "Project Title", "Value"),
    format = "pipe"
  )
```

## Working Papers

- Dai, Yongsheng, Quinn, Barry, Kearney, Fearghal (2025). *Signal Amplification in Market Manipulation Detection*. Working paper.
- Dai, Yongsheng, Quinn, Barry, Kearney, Fearghal, Liu, Weilong, Spence, Ivor, Rafferty, Karen, Wang, Hui (2025). *Detecting Market Manipulation with Self-supervised Learning: A Hybrid Framework Using Frequency-based Synthetic Anomaly and Domain-Specific Features*. Working paper.
- Quinn, Barry (2023). *Explaining AI in Finance: Past, Present, Prospects*. arXiv preprint arXiv:2306.02773.

## PhD Supervision Excellence

**Completed Supervisions (6):**
- Dr. Jiadong Liu (2018): *Momentum in Empirical Asset Pricing*
- Dr. Ashleigh Neil (2019): *Law and Financial Stability*
- Dr. Colm Kelly (2021): *Machine Learning in Empirical Asset Pricing*
- Dr. Qiao Peng (2021): *US Credit Union Mergers: Causes and Consequences*
- Dr. Kevin Johnson (2019): *Industrial Consolidation: The Irish Credit Union Sector*
- Veronica Zhang (2024): *Capital Policy and State Sponsorship in Chinese Banking*

**Current Supervisions (6):** Including students working on AI applications in finance, sustainable investing, and regulatory technology.

# Teaching and Academic Leadership

## Teaching Excellence
- **Programme Director**, MSc Quantitative Finance (2018-2022)
- **Programme Director**, MSc Computational Finance & Trading (2014-2018)
- **Co-Founder**, Queen's Student Managed Fund (2012-2024)
- **Teaching Award**, QUB (2016)

**Current Modules**: Financial Data Science; Causal AI; AI in Trading; Financial Econometrics

## Academic Service
- **Co-founder**, Finance and AI Research Lab (with Dr Fearghal Kearney), Queen's University Belfast (2022-2024)
- **Programme Director**, MSc Quantitative Finance (2018-2022)
- **Lead Developer**, Queen's Business School Remote Analytics Lab Platform
- **Organiser**, Symposium on AI and Future of Financial Regulation (2023)

# Professional Experience

```{r}
employment_academic <- tribble(
  ~Period, ~Position, ~Institution,
  "**2024 - Present**", "Professor of Finance and FinTech", "Ulster University Business School",
  "**2024 - Present**", "Director, Centre for Financial and Regulatory Technology", "Ulster University Business School",
  "**2020 - 2024**", "Senior Lecturer in Finance, Technology and Data Science", "Queen's Management School",
  "**2010 - 2020**", "Lecturer", "Department of Finance, Queen's Management School",
  "**1998 - 2004**", "Currency Trader and Liquidity Manager", "Janus Henderson Investors, London"
) |>
  knitr::kable(
    col.names = c("Period", "Position", "Institution"),
    format = "pipe"
  )
```

# Professional Recognition

- **Chartered Statistician**, Royal Statistical Society (2019)
- **Advanced Data Science Professional**, Alliance of Data Science Professionals (2024)
- **1st Place**, CFA European Quantitative Finance Awards (2018)
- **Associate Research Fellow**, QUB Momentum One Zero (2021-2024)

---
*Full publication list and detailed CV available upon request*
