---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: ../assets/custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 11pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup 
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

- **Address**: Ulster University Business School, Belfast Campus, York Street, Belfast, BT1 5AB, N. Ireland
- **Email**: b.quinn1@ulster.ac.uk
- **GitHub**: [quinfer](https://github.com/quinfer)

# Personal Statement

As a Professor of Finance and FinTech at Ulster University Business School, I bridge the gap between cutting-edge quantitative research and practical financial applications. With 14 peer-reviewed publications (including 1 ABS4* and 11 ABS3 papers), over £2M in research funding, and extensive industry collaboration, I specialize in applying statistical and machine learning techniques to complex problems in financial markets, regulation, and cooperative finance.

My research portfolio spans systemic risk assessment, market manipulation detection, and AI applications in finance, with particular expertise in regulatory technology (RegTech). I have successfully supervised 6 PhD completions to date, with 6 current doctoral students working on frontier topics including generative AI in finance and market anomaly detection.

My unique perspective combines 9 years of financial markets experience as a currency trader and liquidity manager with rigorous academic training. As a Chartered Statistician, I am particularly committed to promoting ethical approaches to data analysis and evidence-based decision-making in an era of increasing algorithmic complexity.

## Education and Professional Qualifications

```{r degrees, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year,
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006",
  "Ph.D. Finance", "Queen's University Belfast", "2012",
  "Chartered Statistician", "Royal Statistical Society", "2019",
  "MSc Artificial Intelligence", "Queen's University Belfast", "Graduating 2025"
) |>
  knitr::kable()
```

# Research Excellence

## Key Publications (Selected from 14 peer-reviewed papers)

**ABS4* Journal:**
- Liu, Weilong, Zhang, Yong, Liu, Kailong, Quinn, Barry, Yang, Xingyu, Peng, Qiao (2024). Evolutionary Multi-Objective Optimisation for Large-Scale Portfolio Selection With Both Random and Uncertain Returns. *IEEE Transactions on Evolutionary Computation* (ABS4)

**Recent ABS3 Publications:**
- McKillop, Donal, Liu, Kailong, Quinn, Barry, Peng, Qiao (Forthcoming). Modelling and Predicting Credit Union Failures. *International Journal of Forecasting* (ABS3)
- Bouri, E., Quinn, B., Sheenan, L. & Tang, Y. (2024). Investigating extreme linkage topology in the aerospace and defence industry. *International Review of Financial Analysis* (ABS3)
- Quinn, Barry, Gallagher, Ronan, Kuosmanen, Timo (2023). Lurking in the shadows: The impact of CO2 emissions target setting on carbon pricing in the Kyoto agreement period. *Energy Economics* (ABS3)

## Research Impact and Funding

```{r}
#| label: tbl-research-grants-academic
#| tbl-cap: "Major Research Grants (Selected)"

grants_academic <- tribble(
  ~Source, ~Role, ~Period, ~Title, ~Amount,
  "UKRI/InnovateUK", "PI", "2024-2025", "Understanding and Enhancing regulatory compliance using AI", "£250K",
  "InnovateUK", "PI", "2021-2023", "Tail Risk Analytics and Stress Testing", "£180K", 
  "InnovateUK", "Co-I", "2017-2019", "AI and Advanced Retail Analytics", "£150K",
  "Various KTPs", "PI/Co-I", "2017-2023", "Multiple Knowledge Transfer Partnerships", "£800K+"
)

grants_academic |>
  knitr::kable(
    col.names = c("Funding Body", "Role", "Period", "Project Title", "Value"),
    format = "pipe"
  )
```

## PhD Supervision Excellence

**Completed Supervisions (6):**
- Dr. Jiadong Liu (2018): *Momentum in Empirical Asset Pricing*
- Dr. Ashleigh Neil (2019): *Law and Financial Stability*
- Dr. Colm Kelly (2021): *Machine Learning in Empirical Asset Pricing*
- Dr. Qiao Peng (2021): *US Credit Union Mergers: Causes and Consequences*
- Dr. Kevin Johnson (2019): *Industrial Consolidation: The Irish Credit Union Sector*
- Veronica Zhang (2024): *Capital Policy and State Sponsorship in Chinese Banking*

**Current Supervisions (6):** Including students working on AI applications in finance, sustainable investing, and regulatory technology.

# Teaching and Academic Leadership

## Teaching Excellence
- **Programme Director**, MSc Quantitative Finance (2018-2022)
- **Programme Director**, MSc Computational Finance & Trading (2014-2018)
- **Co-Founder**, Queen's Student Managed Fund (2012-2024)
- **Teaching Award**, QUB (2016)

## Academic Service
- **Co-director**, Finance and AI Research Lab, Queen's University Belfast (2022-2024)
- **Programme Director**, MSc Quantitative Finance (2018-2022)
- **Lead Developer**, Queen's Business School Remote Analytics Lab Platform
- **Organiser**, Symposium on AI and Future of Financial Regulation (2023)

# Professional Experience

```{r}
employment_academic <- tribble(
  ~Period, ~Position, ~Institution,
  "**2024 - Present**", "Professor of Finance and FinTech", "Ulster University Business School",
  "**2024 - Present**", "Director, Centre for Financial and Regulatory Technology", "Ulster University Business School",
  "**2020 - 2024**", "Senior Lecturer in Finance, Technology and Data Science", "Queen's Management School",
  "**2010 - 2020**", "Lecturer", "Department of Finance, Queen's Management School",
  "**1998 - 2004**", "Currency Trader and Liquidity Manager", "Janus Henderson Investors, London"
) |>
  knitr::kable(
    col.names = c("Period", "Position", "Institution"),
    format = "pipe"
  )
```

# Professional Recognition

- **Chartered Statistician**, Royal Statistical Society (2019)
- **Advanced Data Science Professional**, Alliance of Data Science Professionals (2024)
- **1st Place**, CFA European Quantitative Finance Awards (2018)
- **Associate Research Fellow**, QUB Momentum One Zero (2021-2024)

---
*Full publication list and detailed CV available upon request*
