---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 11pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

{{< include ../sections/_contact.md >}}

# Personal Statement

Professor Barry Quinn is Professor of Finance & Financial Technology at Ulster University Business School and Director of the Centre for Finance and Responsible Technology. The Centre is guided by three pillars: statistical rigour and intellectual humility; practical industry relevance; and responsible algorithmic innovation.

Barry is a Chartered Statistician and views econometrics as an applied statistical science. His work combines applied econometrics and causal inference with trustworthy machine learning and computational approaches to produce decision-relevant evidence for markets, risk, market integrity, and regulatory compliance. In practice, that means making assumptions explicit, quantifying uncertainty, stress-testing results, and being clear about limits—so claims remain proportionate to evidence.

His research spans causal policy evaluation, algorithmic trading risk detection, and trustworthy AI methods. He works collaboratively across finance, statistics, computer science, and law, and teaches across quantitative finance, econometrics, and AI applications in finance, emphasising ethical practice, reproducibility, and critical thinking.

Before entering academia, Barry worked in financial markets, specialising in currency trading and liquidity management. He holds a PhD in Finance and an MSc in Artificial Intelligence (Distinction) from Queen's University Belfast, and works with external partners to ensure research translates into robust practice.

## Education and Professional Qualifications

{{< include ../sections/_education_core.qmd >}}

# Research Excellence

## Research Programme

My research develops across three methodological strands. The first is **trustworthy machine learning for market integrity and risk**, covering representation-learning approaches to market-manipulation detection (Dai, Quinn, Kearney *et al.*, under review at *Information Processing and Management* and at *Economics Letters*) and transformer-based crash-risk modelling (Birem, Abidi Perier, Quinn & Kearney, under review at the *European Journal of Operational Research*). The methodological commitment throughout is that model performance and model trustworthiness are separable properties, both of which must be evidenced before deployment in operational or regulatory settings.

The second is **computer vision for socio-economic measurement**: the Hannon, French, Quinn & O'Hagan geospatial study of vehicle crime in Northern Ireland (submitted to *Insurance: Mathematics and Economics*), and the hierarchical flag-classification programme (4,501-image NI dataset; 70→16→7 taxonomy; ViT-H-14 baseline at 94.78% on the 7-category task; codebase at `economic-flag-classification`). Both treat vision pipelines as measurement instruments rather than accuracy-maximising systems.

The third is **causal and statistical evidence for governance and regulation**: innovation signals and strategic exits in computational approaches to financial regulation (Kearney, Quinn & Pramanick, under review at the *Journal of Business Venturing*); ownership, risk and regulation in Chinese banking (Zhang, Quinn & Sheenan, CRBF WP 25-017); and the collaborative doctoral partnerships with Pytillia, Napier AI and Funds Axis. Across all three strands the methodological priority is that evidence survive operational and regulatory scrutiny — assumptions explicit, uncertainty quantified, and claims proportionate to what the data can support.

## Key Publications (Selected from 14 peer-reviewed papers)

**ABS4 Journal:**

- Liu, Weilong, Zhang, Yong, Liu, Kailong, Quinn, Barry, Yang, Xingyu, Peng, Qiao (2024). Evolutionary Multi-Objective Optimisation for Large-Scale Portfolio Selection With Both Random and Uncertain Returns. *IEEE Transactions on Evolutionary Computation* (ABS4).

**Recent ABS3 Publications:**

- McKillop, Donal, Liu, Kailong, Quinn, Barry, Peng, Qiao (Forthcoming). Modelling and Predicting Credit Union Failures. *International Journal of Forecasting* (ABS3).
- Bouri, E., Quinn, B., Sheenan, L. & Tang, Y. (2024). Investigating extreme linkage topology in the aerospace and defence industry. *International Review of Financial Analysis* (ABS3).
- Quinn, Barry, Gallagher, Ronan, Kuosmanen, Timo (2023). Lurking in the shadows: The impact of CO2 emissions target setting on carbon pricing in the Kyoto agreement period. *Energy Economics* (ABS3).

*Full publication list: see `index.qmd` or request on application.*

## Research Impact and Funding

```{r}
#| label: tbl-grants-academic
#| tbl-cap: "Major Research Grants (Selected)"
#| results: asis
grants_academic <- read.csv("../data/Grant_income.csv", check.names = FALSE) %>%
  filter(Type == "Research") %>%
  arrange(desc(Amount)) %>%
  slice_head(n = 5) %>%
  mutate(
    Period = ifelse(End == "" | is.na(End),
                    as.character(Start),
                    paste0(Start, "–", End)),
    Amount = scales::comma(Amount, prefix = "£")
  ) %>%
  select(Source, `My Role`, Period, Title, Amount)

grants_academic %>%
  knitr::kable(
    col.names = c("Funding Body", "Role", "Period", "Project Title", "Full Economic Costs"),
    format = "pipe"
  )
```

Figures are read from the canonical `data/Grant_income.csv`. For the full grants register and total research income across 13 awards, see `cv_grants.qmd`.

## Working Papers

- Dai, Yongsheng, Quinn, Barry, Kearney, Fearghal, Wang, Hui (Submitted). *Amplifying Market Manipulation Detection Signals*. Submitted to *Economics Letters*. Manuscript EL66561.
- Dai, Yongsheng, Quinn, Barry, Kearney, Fearghal, Liu, Weilong, Spence, Ivor, Rafferty, Karen, Wang, Hui (Submitted). *Detecting Market Manipulation with Dual-branch Self-supervised Learning: A Unified Framework Integrating Frequency-informed Anomaly Synthesis and Domain-Specific Features*. Submitted to *Information Processing and Management*. Manuscript IPM-D-25-06138.
- Hannon, James (Corresponding Author), French, Declan, Quinn, Barry, O'Hagan, Adrian (Submitted). *Geospatial modeling of vehicle crime in Northern Ireland using computer vision to identify environmental factors*. Submitted to *Insurance: Mathematics and Economics*. Manuscript IME-D-25-00419.

## PhD Supervision Excellence

**Completed Supervisions (6):**

{{< include ../sections/_supervision_completed.md >}}

**Current Supervisions (6):** Including students working on AI applications in finance, sustainable investing, and computational approaches to financial regulation — see `cv_grants.qmd` for the full current register.

# Teaching and Academic Leadership

## Teaching Excellence

- **Programme Director**, MSc Quantitative Finance (2018–2022).
- **Programme Director**, MSc Computational Finance & Trading (2014–2018).
- **Co-Founder**, Queen's Student Managed Fund (2012–2024).
- **Teaching Award**, QUB (2016).

**Current Modules**: Financial Data Science; Causal AI; AI in Trading; Financial Econometrics.

## Academic Service

- **Co-founder**, Finance and AI Research Lab (with Dr Fearghal Kearney), Queen's University Belfast (2022–2024).
- **Programme Director**, MSc Quantitative Finance (2018–2022).
- **Lead Developer**, Queen's Business School Remote Analytics Lab Platform.
- **Organiser**, Symposium on AI and the Future of Financial Regulation (2023).
- **Senior Research Fellow** (Sabbatical), International Research Centre on Cooperative Finance, HEC Montreal University, Canada (May–September 2018).
- **Implementation Team Member**, Bloomberg Trading Room, Queen's University Belfast.
- **External Grant Reviewer**, UKFin+ (UKRI specialist funded project) — reviewed feasibility grant application (£100K) for financial services innovation research (2024).

# Professional Experience

```{r}
#| label: tbl-employment-academic
employment_academic <- tribble(
  ~Period, ~Position, ~Institution,
  "**2024 - Present**", "Professor of Finance & Financial Technology", "Ulster University Business School",
  "**2024 - Present**", "Director, Centre for Finance and Responsible Technology", "Ulster University Business School",
  "**2020 - 2024**", "Senior Lecturer in Finance, Technology and Data Science", "Queen's Management School",
  "**2010 - 2020**", "Lecturer", "Department of Finance, Queen's Management School",
  "**1998 - 2004**", "Currency Trader and Liquidity Manager", "Janus Henderson Investors, London"
) |>
  knitr::kable(
    col.names = c("Period", "Position", "Institution"),
    format = "pipe"
  )

employment_academic
```

# Professional Recognition

{{< include ../sections/_professional_recognition.md >}}

- **1st Place**, CFA European Quantitative Finance Awards (2018).
- **Associate Research Fellow**, QUB Momentum One Zero (2021–2024).

---

*Full publication list and detailed CV available upon request.*
