---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 11pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

{{< include ../sections/_contact.md >}}

# Personal Statement

Professor Barry Quinn is Professor of Finance & Financial Technology at Ulster University Business School and Director of the Centre for Finance and Responsible Technology. The Centre is guided by three pillars: statistical rigour and intellectual humility; practical industry relevance; and responsible algorithmic innovation.

Barry is a Chartered Statistician and views econometrics as an applied statistical science. His work combines applied econometrics and causal inference with trustworthy machine learning and computational approaches to produce decision-relevant evidence for markets, risk, market integrity, and regulatory compliance. In practice, that means making assumptions explicit, quantifying uncertainty, stress-testing results, and being clear about limits—so claims remain proportionate to evidence.

His research spans causal policy evaluation, algorithmic trading risk detection, and trustworthy AI methods. He works collaboratively across finance, statistics, computer science, and law, and teaches across quantitative finance, econometrics, and AI applications in finance, emphasising ethical practice, reproducibility, and critical thinking.

Before entering academia, Barry worked in financial markets, specialising in currency trading and liquidity management. He holds a PhD in Finance and an MSc in Artificial Intelligence from Queen's University Belfast, and works with external partners to ensure research translates into robust practice.

## Education and Professional Qualifications

{{< include ../sections/_education_core.qmd >}}

# Research Excellence

## Research Programme

{{< include ../sections/_research_programme.md >}}

## Peer-Reviewed Publications

{{< include ../sections/_peer_reviewed_publications_academic.md >}}

## Research Impact and Funding

```{r}
#| label: tbl-grants-academic
#| tbl-cap: "Major Research Grants (Selected)"
#| results: asis
grants_academic <- read.csv("../data/Grant_income.csv", check.names = FALSE) %>%
  filter(Type == "Research") %>%
  arrange(desc(Amount)) %>%
  slice_head(n = 5) %>%
  mutate(
    Period = ifelse(End == "" | is.na(End),
                    as.character(Start),
                    paste0(Start, "–", End)),
    Amount = scales::comma(Amount, prefix = "£")
  ) %>%
  select(Source, `My Role`, Period, Title, Amount)

grants_academic %>%
  knitr::kable(
    col.names = c("Funding Body", "Role", "Period", "Project Title", "Full Economic Costs"),
    format = "pipe"
  )
```

## Working Papers (Selected)

- Dai, Yongsheng, Quinn, Barry, Kearney, Fearghal, Wang, Hui. *Signal Amplification and Strategic Deterrence in Market Surveillance*. Working paper.
- Dai, Yongsheng, Huang, Ji, Ren, Spence, Ivor, Rafferty, Karen, Quinn, Barry, Wang, Hui. *FMISD: Fine-grained Multi-Scale Modeling for Inexact Supervised Anomaly Detection of Time Series*. Under revision at *IEEE Transactions on Cybernetics*.
- Quinn, Barry, Sheenan, Lisa, Zhang, Ying (Veronica). *Regulatory Reach or Regulatory Arbitrage? Basel III's Cross-Border Transmission Through Foreign Bank Subsidiaries*. Centre for Responsible Banking & Finance Working Paper 25-017.
- Hannon, James (Corresponding Author), French, Declan, Quinn, Barry, O'Hagan, Adrian. *Geospatial modeling of vehicle crime in Northern Ireland using computer vision to identify environmental factors*. Under revision.
- Cochrane, Brandon, Bryan, Dominic, Quinn, Barry, Graham, Bryon, French, Declan (2026). *Flying the Flag: Cultural Expression as Territorial Demarcation*. IGPSJ Working Paper WP 02/26, Queen's University Belfast.

## PhD Supervision Excellence

**Completed Supervisions (6):**

{{< include ../sections/_supervision_completed.md >}}

**Current Supervisions (6):**

{{< include ../sections/_supervision_current.md >}}

# Teaching and Academic Leadership

## Teaching Excellence

- **Programme Director**, MSc Quantitative Finance (2018–2022).
- **Programme Director**, MSc Computational Finance & Trading (2014–2018).
- **Co-Founder**, Queen's Student Managed Fund (2012–2024).
- **Teaching Award**, QUB (2016).

**Current Modules**: Financial Data Science; Causal AI; AI in Trading; Financial Econometrics.

**Teaching software**: [TickLab](https://www.ticklab.co.uk) — web-based trading simulation for market microstructure teaching.

## Academic Service

- **Co-founder**, Finance and AI Research Lab (with Dr Fearghal Kearney), Queen's University Belfast (2022–2024).
- **Programme Director**, MSc Quantitative Finance (2018–2022).
- **Lead Developer**, Queen's Business School Remote Analytics Lab Platform.
- **Organiser**, Symposium on AI and the Future of Financial Regulation (2023).
- **Senior Research Fellow** (Sabbatical), International Research Centre on Cooperative Finance, HEC Montreal University, Canada (May–September 2018).
- **Implementation Team Member**, Bloomberg Trading Room, Queen's University Belfast.
- **External Grant Reviewer**, UKFin+ (UKRI specialist funded project) — reviewed feasibility grant application (£100K) for financial services innovation research (2024).

# Professional Experience

```{r}
#| label: tbl-employment-academic
employment_academic <- tribble(
  ~Period, ~Position, ~Institution,
  "**2024 - Present**", "Professor of Finance & Financial Technology", "Ulster University Business School",
  "**September 2025 - Present**", "Director, Centre for Finance and Responsible Technology", "Ulster University Business School",
  "**2020 - 2024**", "Senior Lecturer in Finance, Technology and Data Science", "Queen's Management School",
  "**2010 - 2020**", "Lecturer", "Department of Finance, Queen's Management School",
  "**1998 - 2004**", "Currency Trader and Liquidity Manager", "Janus Henderson Investors, London"
) |>
  knitr::kable(
    col.names = c("Period", "Position", "Institution"),
    format = "pipe"
  )

employment_academic
```

# Professional Recognition

{{< include ../sections/_professional_recognition.md >}}

- **1st Place**, CFA European Quantitative Finance Awards (2018).
- **Associate Research Fellow**, QUB Momentum One Zero (2021–2024).
