---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: ../assets/custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 11pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup 
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

- **Address**: Ulster University Business School, Belfast Campus, York Street, Belfast, BT1 5AB, N. Ireland
- **Email**: b.quinn1@ulster.ac.uk
- **GitHub**: [quinfer](https://github.com/quinfer)

# Personal Statement

I am Professor of Finance and FinTech at Ulster University Business School and Director of the Centre for Financial and Regulatory Technology. As an applied econometrician and Chartered Statistician, I explore the intersection of finance, technology, and data science, focusing on causal mechanisms beyond cookbook econometrics. My work also engages with computational economics and its finance counterpart, where numerical methods, simulation, and agent-based modelling are applied to problems that defy closed-form solutions.

My research applies statistical, computational, and machine learning techniques to financial markets, causal policy analysis, and cooperative finance. I have been fortunate to receive support from UKRI/EPSRC, Innovate UK, and the Department for the Economy for this work, which has resulted in publications in journals including IEEE Transactions on Evolutionary Computation, International Journal of Forecasting, and Energy Economics.

Before academia, I worked in financial markets as a currency trader and liquidity manager at Janus Henderson Investors, an experience that continues to inform my approach to financial research. I hold a PhD in Finance from Queen's University Belfast and am currently pursuing an MSc in Artificial Intelligence to better understand emerging technologies in finance.

Through my teaching in quantitative finance, econometrics, financial data science, causal AI, and AI in trading, I hope to help students develop both technical competence and ethical approaches to data analysis. By framing my work within computational finance as well as econometrics, I aim to contribute meaningfully to evidence-based decision-making in finance whilst maintaining the rigorous statistical standards expected in our field.

## Education and Professional Qualifications

```{r degrees, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year,
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006",
  "Ph.D. Finance", "Queen's University Belfast", "2012",
  "Chartered Statistician", "Royal Statistical Society", "2019",
  "MSc Artificial Intelligence", "Queen's University Belfast", "Graduating 2025"
) |>
  knitr::kable()
```

# Research Excellence

## Key Publications (Selected from 14 peer-reviewed papers)

**ABS4* Journal:**
- Liu, Weilong, Zhang, Yong, Liu, Kailong, Quinn, Barry, Yang, Xingyu, Peng, Qiao (2024). Evolutionary Multi-Objective Optimisation for Large-Scale Portfolio Selection With Both Random and Uncertain Returns. *IEEE Transactions on Evolutionary Computation* (ABS4)

**Recent ABS3 Publications:**
- McKillop, Donal, Liu, Kailong, Quinn, Barry, Peng, Qiao (Forthcoming). Modelling and Predicting Credit Union Failures. *International Journal of Forecasting* (ABS3)
- Bouri, E., Quinn, B., Sheenan, L. & Tang, Y. (2024). Investigating extreme linkage topology in the aerospace and defence industry. *International Review of Financial Analysis* (ABS3)
- Quinn, Barry, Gallagher, Ronan, Kuosmanen, Timo (2023). Lurking in the shadows: The impact of CO2 emissions target setting on carbon pricing in the Kyoto agreement period. *Energy Economics* (ABS3)

## Research Impact and Funding

```{r}
#| label: tbl-research-grants-academic
#| tbl-cap: "Major Research Grants (Selected)"

grants_academic <- tribble(
  ~Source, ~Role, ~Period, ~Title, ~Amount,
  "UKRI/InnovateUK", "PI", "2024-2025", "Understanding and Enhancing regulatory compliance using AI", "£250K",
  "InnovateUK", "PI", "2021-2023", "Tail Risk Analytics and Stress Testing", "£180K", 
  "InnovateUK", "Co-I", "2017-2019", "AI and Advanced Retail Analytics", "£150K",
  "Various KTPs", "PI/Co-I", "2017-2023", "Multiple Knowledge Transfer Partnerships", "£800K+"
)

grants_academic |>
  knitr::kable(
    col.names = c("Funding Body", "Role", "Period", "Project Title", "Value"),
    format = "pipe"
  )
```

## Working Papers

- Dai, Yongsheng, Quinn, Barry, Kearney, Fearghal (2025). *Signal Amplification in Market Manipulation Detection*. Working paper.
- Dai, Yongsheng, Quinn, Barry, Kearney, Fearghal, Liu, Weilong, Spence, Ivor, Rafferty, Karen, Wang, Hui (2025). *Detecting Market Manipulation with Self-supervised Learning: A Hybrid Framework Using Frequency-based Synthetic Anomaly and Domain-Specific Features*. Working paper.
- Quinn, Barry (2023). *Explaining AI in Finance: Past, Present, Prospects*. arXiv preprint arXiv:2306.02773.

## PhD Supervision Excellence

**Completed Supervisions (6):**
- Dr. Jiadong Liu (2018): *Momentum in Empirical Asset Pricing*
- Dr. Ashleigh Neil (2019): *Law and Financial Stability*
- Dr. Colm Kelly (2021): *Machine Learning in Empirical Asset Pricing*
- Dr. Qiao Peng (2021): *US Credit Union Mergers: Causes and Consequences*
- Dr. Kevin Johnson (2019): *Industrial Consolidation: The Irish Credit Union Sector*
- Veronica Zhang (2024): *Capital Policy and State Sponsorship in Chinese Banking*

**Current Supervisions (6):** Including students working on AI applications in finance, sustainable investing, and regulatory technology.

# Teaching and Academic Leadership

## Teaching Excellence
- **Programme Director**, MSc Quantitative Finance (2018-2022)
- **Programme Director**, MSc Computational Finance & Trading (2014-2018)
- **Co-Founder**, Queen's Student Managed Fund (2012-2024)
- **Teaching Award**, QUB (2016)

**Current Modules**: Financial Data Science; Causal AI; AI in Trading; Financial Econometrics

## Academic Service
- **Co-director**, Finance and AI Research Lab, Queen's University Belfast (2022-2024)
- **Programme Director**, MSc Quantitative Finance (2018-2022)
- **Lead Developer**, Queen's Business School Remote Analytics Lab Platform
- **Organiser**, Symposium on AI and Future of Financial Regulation (2023)

# Professional Experience

```{r}
employment_academic <- tribble(
  ~Period, ~Position, ~Institution,
  "**2024 - Present**", "Professor of Finance and FinTech", "Ulster University Business School",
  "**2024 - Present**", "Director, Centre for Financial and Regulatory Technology", "Ulster University Business School",
  "**2020 - 2024**", "Senior Lecturer in Finance, Technology and Data Science", "Queen's Management School",
  "**2010 - 2020**", "Lecturer", "Department of Finance, Queen's Management School",
  "**1998 - 2004**", "Currency Trader and Liquidity Manager", "Janus Henderson Investors, London"
) |>
  knitr::kable(
    col.names = c("Period", "Position", "Institution"),
    format = "pipe"
  )
```

# Professional Recognition

- **Chartered Statistician**, Royal Statistical Society (2019)
- **Advanced Data Science Professional**, Alliance of Data Science Professionals (2024)
- **1st Place**, CFA European Quantitative Finance Awards (2018)
- **Associate Research Fellow**, QUB Momentum One Zero (2021-2024)

---
*Full publication list and detailed CV available upon request*
