---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: ../assets/custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 10pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup 
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

- **Address**: Ulster University Business School, Belfast Campus, York Street, Belfast, BT1 5AB, N. Ireland
- **Email**: b.quinn1@ulster.ac.uk | **GitHub**: [quinfer](https://github.com/quinfer)
- **ORCID**: [0000-0002-8637-9060](https://orcid.org/0000-0002-8637-9060)
- **Date of Birth**: 10 July 1973

# Research Profile Summary

Professor Barry Quinn is Professor of Finance & Financial Technology at Ulster University Business School, where he is Director of the Centre for Finance and Responsible Technology. The Centre has been established with £360K in PhD scholarships from the Department of the Economy NI, including collaborative doctoral partnerships with local fintech firms Pytillia and Napier AI. The Centre is built on three pillars of excellence: statistical rigour and intellectual humility, practical industry relevance, and responsible algorithmic innovation.

As an applied econometrician and Chartered Statistician, Barry's work integrates econometrics, machine learning, and computational approaches to address challenges in financial markets, risk, and regulatory compliance. Financial decisions have real consequences, and his approach emphasises making assumptions explicit, quantifying uncertainty, and stress-testing results so we can say plainly when evidence does (and does not) support confident conclusions. He values rigorous peer review and publication in high-quality journals, and also asks whether work is transparent, reproducible, and useful to those making decisions.

His research addresses real financial challenges through the co-development of research and innovation with technology companies and financial institutions. This practical industry relevance ensures that academic research translates into solutions that matter. Barry's work combines causal policy analysis, algorithmic trading risk detection, and trustworthy AI methods, with a particular interest in how financial technology can support resilient markets and evidence-based regulation. His research is collaborative and interdisciplinary, bridging economics, statistics, computer science, and law.

Responsible algorithmic innovation starts with the same discipline: clarity about assumptions, evidence, and limits. Barry develops and evaluates algorithmic systems that are transparent and auditable, tested under realistic conditions. This includes attention to explainability, bias and fairness risks, validation and controls, and privacy, so that innovation improves financial services whilst maintaining public trust and regulatory compliance.

Before entering academia, Barry worked in financial markets, specialising in currency trading and liquidity management. He holds a PhD in Finance from Queen's University Belfast, is a Chartered Statistician, and holds an MSc in Artificial Intelligence with Distinction from Queen's University Belfast. He actively collaborates with industry and policy partners to ensure his research delivers academic, societal, and regulatory impact.

**Research Interests**: AI applications in finance, causal AI, financial data science, market manipulation detection, systemic risk assessment, computational approaches to financial regulation, cooperative finance, sustainable finance, generative AI in capital markets.

# Education and Professional Qualifications

```{r degrees, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year, ~Relevance,
  "Ph.D. Finance", "Queen's University Belfast", "2012", "Thesis: Cooperative Financial Institutions",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006", "Advanced derivatives and risk management",
  "MSc Artificial Intelligence (Distinction)", "Queen's University Belfast", "2025", "Deep learning, NLP, computer vision",
  "Chartered Statistician", "Royal Statistical Society", "2019", "Professional statistical expertise",
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995", "First Class Honours"
) |>
  knitr::kable(
    col.names = c("Qualification", "Institution", "Year", "Research Relevance"),
    format = "pipe"
  )
```

# Research Excellence

## Publication Record (h-index: Available upon request)

### Peer-Reviewed Publications (14 total)

**Recent High-Impact Publications:**

- **Liu, Weilong, Zhang, Yong, Liu, Kailong, Quinn, Barry**, Yang, Xingyu, Peng, Qiao (2024). Evolutionary Multi-Objective Optimisation for Large-Scale Portfolio Selection With Both Random and Uncertain Returns. *IEEE Transactions on Evolutionary Computation* (**ABS4**)

- **McKillop, Donal, Liu, Kailong, Quinn, Barry**, Peng, Qiao (Forthcoming). Modelling and Predicting Credit Union Failures. *International Journal of Forecasting* (**ABS3**)

- **Bouri, E., Quinn, B., Sheenan, L. & Tang, Y.** (2024). Investigating extreme linkage topology in the aerospace and defence industry. *International Review of Financial Analysis* (**ABS3**)

- **Quinn, Barry, Gallagher, Ronan, Kuosmanen, Timo** (2023). Lurking in the shadows: The impact of CO2 emissions target setting on carbon pricing in the Kyoto agreement period. *Energy Economics* (**ABS3**)

### Software Publications
- **Time Series Econometrics (tsfe)** (2022) - R Package for advanced time series analysis (DOI: 10.5281/zenodo.6376113)
- **Financial Machine Learning (fml)** (2020) - R Package for financial ML applications (DOI: 10.5281/zenodo.6376114)

### Working Papers & Under Review
- **Kearney, Fearghal, Quinn, Barry, Dai, Yongsheng** (2025) *Signal Amplification in Market Manipulation Detection*
- **Dai, Yongsheng, Quinn, Barry, Kearney, Fearghal, Liu, Weilong, Spence, Ivor, Rafferty, Karen, Wang, Hui** (2025). *Detecting Market Manipulation with Self-supervised Learning: A Hybrid Framework Using Frequency-based Synthetic Anomaly and Domain-Specific Features*. Working paper.
- **Quinn, Barry** (2023). *Explaining AI in Finance: Past, Present, Prospects*. arXiv preprint arXiv:2306.02773.
- **Zhang, Ying (Veronica), Quinn, Barry, Sheenan, Lisa** (2025). *Ownership Dynamics, Risk and Regulation in Chinese Banking: New Evidence*
- **Kearney, Fearghal, Quinn, Barry, Pramanick, Abhishek** (Under Review) *Innovation Signals and Strategic Exits: How Technological Readiness Shapes Outcomes in Computational Approaches to Financial Regulation*

# Research Funding Track Record

```{r}
#| label: tbl-grants-detailed
#| tbl-cap: "Research Grant Portfolio"

# Simple test - just create a basic table with hardcoded values
test_data <- data.frame(
  Source = c("Department of the Economy NI", "Department of the Economy NI", "Department of the Economy NI", "Innovate UK", "Innovate UK", "UKRI/InnovateUK", "DofE CAST Award NI"),
  Role = c("PI", "PI", "PI", "PI", "PI", "PI", "Co-I"),
  Period = c("2025", "2025", "2025", "2022-2023", "2021-2023", "2024-2025", "2023"),
  Title = c("PhD Scholarship - Collaborative Doctoral Partnership with Pytillia", "PhD Scholarship - Collaborative Doctoral Partnership with Napier AI", "PhD Scholarship - Centre for Finance and Responsible Technology", "E.S.G fair value analytics platform", "Computational approaches to financial regulation and portfolio analytics", "AI regulatory compliance in Investment Management", "Welfare Cost of Cultural Displays using AI"),
  Value = c("£120,000", "£120,000", "£120,000", "£173,000", "£173,000", "£100,000", "£70,000"),
  Type = c("PhD Scholarship", "PhD Scholarship", "PhD Scholarship", "Research", "Research", "Research", "Research")
)

knitr::kable(test_data, 
             col.names = c("Funding Body", "Role", "Period", "Project Title", "Value", "Type"),
             format = "pipe")
```

**Total Research Income**: £1,435,000+ (direct awards as PI/Co-I)  
**Success Rate**: High success rate in competitive funding schemes  
**Industry Leverage**: Strong track record of industry co-funding and partnerships, including collaborative doctoral partnerships with fintech firms

# Research Impact and Knowledge Transfer

## Industry Collaborations
- **Pytillia** - Collaborative Doctoral Partnership (PhD scholarship, £120K, 2025)
- **Napier AI** - Collaborative Doctoral Partnership (PhD scholarship, £120K, 2025)
- **Funds Axis Ltd** - Computational approaches to financial regulation and compliance solutions (ongoing)
- **Citigroup Belfast** - Market manipulation detection (2022-2023)
- **Pearl AI Ltd** - Advanced retail analytics (2017-2019)
- **Financial Conduct Authority** - Tech Sprint Mentor (2023-present)

## Research Supervision Excellence

### Completed PhD Supervisions (6)
1. **Dr. Jiadong Liu** (2018): *Momentum in Empirical Asset Pricing* - Now research scientist
2. **Dr. Ashleigh Neil** (2019): *Law and Financial Stability* - Now policy analyst
3. **Dr. Colm Kelly** (2021): *Machine Learning in Empirical Asset Pricing* - Now quant researcher
4. **Dr. Qiao Peng** (2021): *US Credit Union Mergers* - Now assistant professor
5. **Dr. Kevin Johnson** (2019): *Industrial Consolidation* - Now senior analyst
6. **Veronica Zhang** (2024): *Capital Policy in Chinese Banking* - Graduating

### Current PhD Supervisions (6)
- **Fei Li**: Sustainable Investing and Gender Diversity (QBS scholarship)
- **Abhishek Pramanick**: AI and Future of Financial Regulation (QBS scholarship)  
- **Dongwei Li**: Economics of AI Patents in Finance (DofE scholarship)
- **Yuqi Ding**: Economics of Generative AI in Finance (QBS scholarship)
- **Yongsheng Dai**: Market Manipulation Detection (PWC scholarship)
- **Brandon Cochrane**: Economic Cost of Cultural Displays (DofE CAST scholarship)

## International Research Network
- **Visiting PhD Supervisors**: Students from China (SYSU), Ireland (UCD), France (UPEC Paris)
- **International Collaborations**: Active research partnerships across Europe, Asia, and North America
- **Conference Presentations**: Regular invited speaker at major international conferences

# Research Infrastructure and Capabilities

## Technical Expertise
- **Advanced Statistical Methods**: Time series econometrics, machine learning, deep learning
- **Programming Languages**: R (expert), Python (advanced), SQL, MATLAB
- **Specialized Software**: Bloomberg Terminal, Reuters, specialized financial databases
- **High-Performance Computing**: Experience with cluster computing and big data analytics

## Research Facilities
- **Queen's Business School Remote Analytics Lab Platform** (Lead Developer, 2021-2024)
- **Finance and AI Research Lab** (Co-director, 2022-2024)
- **Ulster University Centre for Finance and Responsible Technology** (Director, 2024-present)

## Training & Capacity Building
- Design and delivery of graduate teaching in Financial Data Science and causal AI, with applications in trading, market microstructure, and policy analysis
- Supervision and mentoring across econometrics, machine learning, and computational finance, including industry-partnered projects
- Emphasis on reproducible research (versioned data pipelines, model governance) and rigorous statistical standards for evidence-based decision-making

# Future Research Directions

## Emerging Research Themes
1. **Generative AI in Capital Markets**: Applications and regulatory implications
2. **Explainable AI for Financial Regulation**: Transparent algorithmic decision-making
3. **Systemic Risk in Digital Finance**: Cryptocurrency and DeFi risk assessment
4. **Sustainable Finance Analytics**: ESG integration and climate risk modeling

## Methodology Innovation
- Integration of large language models with traditional econometric approaches
- Development of hybrid AI systems for financial decision-making
- Advanced causal inference methods for policy evaluation

# Professional Recognition and Service

## Academic Leadership
- **Director**, Centre for Finance and Responsible Technology (2024-present)
- **Programme Director**, MSc Quantitative Finance (2018-2022)
- **Organizer**, Major symposiums on AI and Financial Regulation

## Professional Recognition
- **Chartered Statistician**, Royal Statistical Society (2019)
- **1st Place**, CFA European Quantitative Finance Awards (2018)
- **Advanced Data Science Professional**, Royal Statistical Society (2024)

## Editorial and Review Activities
- Regular reviewer for major finance and technology journals
- Conference program committee member
- Grant review panels (available upon request)

---

**Research Philosophy**: Bridging cutting-edge quantitative methods with practical financial applications, maintaining statistical rigor while ensuring real-world relevance and regulatory compliance.
