---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 10pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

{{< include ../sections/_contact.md >}}

# Research Profile Summary

Professor Barry Quinn is Professor of Finance & Financial Technology at Ulster University Business School and Director of the Centre for Finance and Responsible Technology. The Centre is guided by three pillars: statistical rigour and intellectual humility; practical industry relevance; and responsible algorithmic innovation.

Barry is a Chartered Statistician and views econometrics as an applied statistical science. His work combines applied econometrics and causal inference with trustworthy machine learning and computational approaches to address problems in financial markets and institutions, risk measurement, market integrity, and evidence for regulation and compliance. A core aim is to produce analysis that is transparent, reproducible, and auditable, and to keep conclusions proportionate to the strength of the evidence.

Before entering academia, Barry worked in financial markets, specialising in currency trading and liquidity management. He holds a PhD in Finance and an MSc in Artificial Intelligence (Distinction) from Queen's University Belfast. He actively collaborates with industry and policy partners to ensure his research delivers academic, societal, and public value.

**Research Interests**: AI applications in finance, causal AI, financial data science, market manipulation detection, systemic risk assessment, computational approaches to financial regulation, cooperative finance, sustainable finance, generative AI in capital markets.

# Education and Professional Qualifications

```{r degrees-grants, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year, ~Relevance,
  "Ph.D. Finance", "Queen's University Belfast", "2012", "Thesis: Cooperative Financial Institutions",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006", "Advanced derivatives and risk management",
  "MSc Artificial Intelligence (Distinction)", "Queen's University Belfast", "2025", "Deep learning, NLP, computer vision",
  "Advanced Data Science Professional", "Royal Statistical Society", "2023", "Professional recognition in data science",
  "Chartered Statistician", "Royal Statistical Society", "2019", "Professional statistical expertise",
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995", "First Class Honours"
) |>
  knitr::kable(
    col.names = c("Qualification", "Institution", "Year", "Research Relevance"),
    format = "pipe"
  )
```

# Research Excellence

## Publication Record

Fourteen peer-reviewed publications (1 ABS4, 11 ABS3s, 2 ABS2s), three commissioned research reports, and two statistical software packages.

### Peer-Reviewed Publications

{{< include ../sections/_peer_reviewed_publications.md >}}

### Software Publications

{{< include ../sections/_software_publications.qmd >}}

### Working Papers and Under Review

{{< include ../sections/_working_papers.md >}}

# Research Funding Track Record

{{< include ../sections/_grants_table.qmd >}}

{{< include ../sections/_grants_total.qmd >}}

**Industry Leverage**: Strong track record of industry co-funding and partnerships, including collaborative doctoral partnerships with local firms.  

# Research Impact and Knowledge Transfer

{{< include ../sections/_impact_projects.md >}}

## Industry Collaborations

- **Pytillia** — Collaborative Doctoral Partnership (PhD scholarship, £120K, 2025).
- **Napier AI** — Collaborative Doctoral Partnership (PhD scholarship, £120K, 2025).
- **Funds Axis Ltd** — Computational approaches to financial regulation and compliance solutions (ongoing).
- **Citigroup Belfast** — Market manipulation detection (2022–2023).
- **Pearl AI Ltd** — Advanced retail analytics (2017–2019).
- **Financial Conduct Authority** — Tech Sprint Mentor (2023–present).

# Research Supervision

## Completed PhD Supervisions

{{< include ../sections/_supervision_completed.md >}}

## Current PhD Supervisions

{{< include ../sections/_supervision_current.md >}}

## Visiting PhD Supervision

{{< include ../sections/_supervision_visiting.md >}}

## International Research Network

- **Visiting PhD Supervisors**: Students from China (SYSU), Ireland (UCD), France (UPEC Paris).
- **International Collaborations**: Active research partnerships across Europe, Asia, and North America.
- **Conference Presentations**: Regular invited speaker at major international conferences — see below.

## Conference Presentations

{{< include ../sections/_conference_talks.md >}}

# Research Infrastructure and Capabilities

## Technical Expertise

- **Advanced statistical methods**: time series econometrics, machine learning, deep learning.
- **Programming languages**: R (expert), Python (advanced), SQL, MATLAB.
- **Specialised software**: Bloomberg Terminal, Reuters, specialised financial databases.
- **High-performance computing**: cluster computing and big-data analytics.

## Research Facilities

- **Queen's Business School Remote Analytics Lab Platform** (Lead Developer, 2021–2024).
- **Finance and AI Research Lab** (Co-director, 2022–2024).
- **Ulster University Centre for Finance and Responsible Technology** (Director, 2024–present).

## Training and Capacity Building

- Design and delivery of graduate teaching in Financial Data Science and causal AI, with applications in trading, market microstructure, and policy analysis.
- Supervision and mentoring across econometrics, machine learning, and computational finance, including industry-partnered projects.
- Emphasis on reproducible research (versioned data pipelines, model governance) and rigorous statistical standards for evidence-based decision-making.

# Future Research Directions

## Emerging Research Themes

1. **Generative AI in capital markets**: applications and regulatory implications.
2. **Explainable AI for financial regulation**: transparent algorithmic decision-making.
3. **Systemic risk in digital finance**: cryptocurrency and DeFi risk assessment.
4. **Sustainable finance analytics**: ESG integration and climate risk modelling.

## Methodology Innovation

- Integration of large language models with traditional econometric approaches.
- Development of hybrid AI systems for financial decision-making.
- Advanced causal inference methods for policy evaluation.

# Professional Recognition and Service

## Academic Leadership

{{< include ../sections/_academic_service.md >}}

## Professional Recognition

{{< include ../sections/_professional_recognition.md >}}

## Awards

{{< include ../sections/_awards.md >}}

---

**Research Philosophy**: Bridging cutting-edge quantitative methods with practical financial applications, maintaining statistical rigour while ensuring real-world relevance and regulatory compliance.
