---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: ../assets/custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 10pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup 
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

- **Address**: Ulster University Business School, Belfast Campus, York Street, Belfast, BT1 5AB, N. Ireland
- **Email**: b.quinn1@ulster.ac.uk | **GitHub**: [quinfer](https://github.com/quinfer)

# Research Profile Summary

Professor of Finance and FinTech at Ulster University Business School and Director of the Centre for Financial and Regulatory Technology. As an applied econometrician and Chartered Statistician, my work sits at the intersection of finance, technology, and data science with an emphasis on causal mechanisms beyond cookbook econometrics. I also engage with computational economics and its finance counterpart, applying numerical methods, simulation, and agent-based modelling to problems that defy closed-form solutions.

My research applies statistical, computational, and machine learning techniques to financial markets, causal policy analysis, and cooperative finance. This portfolio has been supported by UKRI/EPSRC, Innovate UK, and the Department for the Economy and has produced publications in venues including IEEE Transactions on Evolutionary Computation, International Journal of Forecasting, and Energy Economics.

Prior to academia I worked in financial markets as a currency trader and liquidity manager at Janus Henderson Investors. I hold a PhD in Finance from Queen's University Belfast and am currently pursuing an MSc in Artificial Intelligence. These experiences support the delivery of rigorous, policy-relevant, and translational research with industry and public sector partners.

**Research Interests**: AI applications in finance, causal AI, financial data science, market manipulation detection, systemic risk assessment, regulatory technology, cooperative finance, sustainable finance, generative AI in capital markets.

# Education and Professional Qualifications

```{r degrees, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year, ~Relevance,
  "Ph.D. Finance", "Queen's University Belfast", "2012", "Thesis: Cooperative Financial Institutions",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006", "Advanced derivatives and risk management",
  "MSc Artificial Intelligence", "Queen's University Belfast", "2025*", "Current: Deep learning, NLP, computer vision",
  "Chartered Statistician", "Royal Statistical Society", "2019", "Professional statistical expertise",
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995", "First Class Honours"
) |>
  knitr::kable(
    col.names = c("Qualification", "Institution", "Year", "Research Relevance"),
    format = "pipe"
  )
```

# Research Excellence

## Publication Record (h-index: Available upon request)

### Peer-Reviewed Publications (14 total)

**Recent High-Impact Publications:**

- **Liu, Weilong, Zhang, Yong, Liu, Kailong, Quinn, Barry**, Yang, Xingyu, Peng, Qiao (2024). Evolutionary Multi-Objective Optimisation for Large-Scale Portfolio Selection With Both Random and Uncertain Returns. *IEEE Transactions on Evolutionary Computation* (**ABS4**)

- **McKillop, Donal, Liu, Kailong, Quinn, Barry**, Peng, Qiao (Forthcoming). Modelling and Predicting Credit Union Failures. *International Journal of Forecasting* (**ABS3**)

- **Bouri, E., Quinn, B., Sheenan, L. & Tang, Y.** (2024). Investigating extreme linkage topology in the aerospace and defence industry. *International Review of Financial Analysis* (**ABS3**)

- **Quinn, Barry, Gallagher, Ronan, Kuosmanen, Timo** (2023). Lurking in the shadows: The impact of CO2 emissions target setting on carbon pricing in the Kyoto agreement period. *Energy Economics* (**ABS3**)

### Software Publications
- **Time Series Econometrics (tsfe)** (2022) - R Package for advanced time series analysis (DOI: 10.5281/zenodo.6376113)
- **Financial Machine Learning (fml)** (2020) - R Package for financial ML applications (DOI: 10.5281/zenodo.6376114)

### Working Papers & Under Review
- **Kearney, Fearghal, Quinn, Barry, Dai, Yongsheng** (2025) *Signal Amplification in Market Manipulation Detection*
- **Zhang, Ying (Veronica), Quinn, Barry, Sheenan, Lisa** (2025). *Ownership Dynamics, Risk and Regulation in Chinese Banking: New Evidence*
- **Kearney, Fearghal, Quinn, Barry, Pramanick, Abhishek** (Under Review) *Innovation Signals and Strategic Exits: How Technological Readiness Shapes RegTech Venture Outcomes*

# Research Funding Track Record

```{r}
#| label: tbl-grants-detailed
#| tbl-cap: "Research Grant Portfolio"

# Simple test - just create a basic table with hardcoded values
test_data <- data.frame(
  Source = c("Innovate UK", "Innovate UK", "UKRI/InnovateUK", "DofE CAST Award NI"),
  Role = c("PI", "PI", "PI", "Co-I"),
  Period = c("2022-2023", "2021-2023", "2024-2025", "2023"),
  Title = c("E.S.G fair value analytics platform", "Regulatory technology and portfolio analytics", "AI regulatory compliance in Investment Management", "Welfare Cost of Cultural Displays using AI"),
  Value = c("£173,000", "£173,000", "£100,000", "£70,000"),
  Type = c("Research", "Research", "Research", "Research")
)

knitr::kable(test_data, 
             col.names = c("Funding Body", "Role", "Period", "Project Title", "Value", "Type"),
             format = "pipe")
```

**Total Research Income**: £1,075,000+ (direct awards as PI/Co-I)  
**Success Rate**: High success rate in competitive funding schemes  
**Industry Leverage**: Strong track record of industry co-funding and partnerships

# Research Impact and Knowledge Transfer

## Industry Collaborations
- **Funds Axis Ltd** - RegTech and compliance solutions (ongoing)
- **Citigroup Belfast** - Market manipulation detection (2022-2023)
- **Pearl AI Ltd** - Advanced retail analytics (2017-2019)
- **Financial Conduct Authority** - Tech Sprint Mentor (2023-present)

## Research Supervision Excellence

### Completed PhD Supervisions (6)
1. **Dr. Jiadong Liu** (2018): *Momentum in Empirical Asset Pricing* - Now research scientist
2. **Dr. Ashleigh Neil** (2019): *Law and Financial Stability* - Now policy analyst
3. **Dr. Colm Kelly** (2021): *Machine Learning in Empirical Asset Pricing* - Now quant researcher
4. **Dr. Qiao Peng** (2021): *US Credit Union Mergers* - Now assistant professor
5. **Dr. Kevin Johnson** (2019): *Industrial Consolidation* - Now senior analyst
6. **Veronica Zhang** (2024): *Capital Policy in Chinese Banking* - Graduating

### Current PhD Supervisions (6)
- **Fei Li**: Sustainable Investing and Gender Diversity (QBS scholarship)
- **Abhishek Pramanick**: AI and Future of Financial Regulation (QBS scholarship)  
- **Dongwei Li**: Economics of AI Patents in Finance (DofE scholarship)
- **Yuqi Ding**: Economics of Generative AI in Finance (QBS scholarship)
- **Yongsheng Dai**: Market Manipulation Detection (PWC scholarship)
- **Brandon Cochrane**: Economic Cost of Cultural Displays (DofE CAST scholarship)

## International Research Network
- **Visiting PhD Supervisors**: Students from China (SYSU), Ireland (UCD), France (UPEC Paris)
- **International Collaborations**: Active research partnerships across Europe, Asia, and North America
- **Conference Presentations**: Regular invited speaker at major international conferences

# Research Infrastructure and Capabilities

## Technical Expertise
- **Advanced Statistical Methods**: Time series econometrics, machine learning, deep learning
- **Programming Languages**: R (expert), Python (advanced), SQL, MATLAB
- **Specialized Software**: Bloomberg Terminal, Reuters, specialized financial databases
- **High-Performance Computing**: Experience with cluster computing and big data analytics

## Research Facilities
- **Queen's Business School Remote Analytics Lab Platform** (Lead Developer, 2021-2024)
- **Finance and AI Research Lab** (Co-director, 2022-2024)
- **Ulster University Centre for Financial and Regulatory Technology** (Director, 2024-present)

## Training & Capacity Building
- Design and delivery of graduate teaching in Financial Data Science and causal AI, with applications in trading, market microstructure, and policy analysis
- Supervision and mentoring across econometrics, machine learning, and computational finance, including industry-partnered projects
- Emphasis on reproducible research (versioned data pipelines, model governance) and rigorous statistical standards for evidence-based decision-making

# Future Research Directions

## Emerging Research Themes
1. **Generative AI in Capital Markets**: Applications and regulatory implications
2. **Explainable AI for Financial Regulation**: Transparent algorithmic decision-making
3. **Systemic Risk in Digital Finance**: Cryptocurrency and DeFi risk assessment
4. **Sustainable Finance Analytics**: ESG integration and climate risk modeling

## Methodology Innovation
- Integration of large language models with traditional econometric approaches
- Development of hybrid AI systems for financial decision-making
- Advanced causal inference methods for policy evaluation

# Professional Recognition and Service

## Academic Leadership
- **Director**, Centre for Financial and Regulatory Technology (2024-present)
- **Programme Director**, MSc Quantitative Finance (2018-2022)
- **Organizer**, Major symposiums on AI and Financial Regulation

## Professional Recognition
- **Chartered Statistician**, Royal Statistical Society (2019)
- **1st Place**, CFA European Quantitative Finance Awards (2018)
- **Advanced Data Science Professional** (2024)

## Editorial and Review Activities
- Regular reviewer for major finance and technology journals
- Conference program committee member
- Grant review panels (available upon request)

---

**Research Philosophy**: Bridging cutting-edge quantitative methods with practical financial applications, maintaining statistical rigor while ensuring real-world relevance and regulatory compliance.
