---
title: "Prof. Barry Quinn"
subtitle: "Finance & FinTech Leader | AI & Data Science Expert"
format:
  html:
    template: ../assets/custom-template.html
    css: ../assets/custom-styles.css
  pdf:
    documentclass: article
    geometry:
      - margin=0.75in
    fontsize: 11pt
    linkcolor: blue
    urlcolor: blue
---

```{r}
#| label: setup 
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

**Prof. Barry Quinn** | Professor of Finance and FinTech  
📧 b.quinn1@ulster.ac.uk | 🔗 [GitHub: quinfer](https://github.com/quinfer)  
📍 Ulster University Business School, Belfast, N. Ireland

---

# Executive Summary

Applied econometrician and Chartered Statistician working at the intersection of finance, technology, and data science. I focus on causal mechanisms beyond cookbook econometrics and on computational finance approaches—numerical methods, simulation, and agent-based modelling—to solve problems that resist closed-form analysis. My research and partnerships apply statistical, computational, and machine learning techniques to financial markets, regulatory compliance, and policy, with current teaching and delivery in Financial Data Science and causal AI. This work has been supported by UKRI/EPSRC, Innovate UK, and the Department for the Economy, and has resulted in publications in IEEE Transactions on Evolutionary Computation, International Journal of Forecasting, and Energy Economics. Prior to academia, I worked as a currency trader and liquidity manager at Janus Henderson Investors; that experience underpins a pragmatic approach to building AI- and analytics-driven solutions. I hold a PhD in Finance and am currently completing an MSc in Artificial Intelligence.

**Core Value Proposition:**
- **Industry Experience**: Former currency trader and liquidity manager at Janus Henderson Investors
- **Academic Excellence**: 14 peer-reviewed publications, Chartered Statistician
- **Technology Leadership**: AI/ML expert currently completing MSc in Artificial Intelligence
- **Business Impact**: £2M+ research funding, multiple successful industry partnerships

---

# Key Expertise & Skills

## Technical Competencies
- **AI & Machine Learning**: Deep learning, causal AI, market manipulation detection, anomaly detection
- **Financial Data Science**: Data pipelines, feature engineering, time-series ML, MLOps
- **Financial Technology**: RegTech, algorithmic trading, risk analytics, stress testing
- **Data Science**: Advanced statistical modeling, big data analytics, predictive modeling
- **Programming**: R, Python, statistical software development

## Business Applications
- **Risk Management**: Systemic risk assessment, VaR modeling, stress testing
- **Regulatory Compliance**: AI-driven compliance solutions, regulatory technology
- **Trading & Markets**: Currency trading, liquidity management, market microstructure
- **Strategic Analytics**: ESG analytics, performance measurement, scenario analysis

---

# Professional Experience

```{r}
employment_industry <- tribble(
  ~Period, ~Role, ~Organization, ~Key_Achievements,
  "2024-Present", "Professor of Finance & FinTech", "Ulster University", "Director of Centre for Financial & Regulatory Technology",
  "2020-2024", "Senior Lecturer", "Queen's University Belfast", "Led £2M+ research portfolio, 12+ industry partnerships",
  "2010-2020", "Lecturer & Programme Director", "Queen's University Belfast", "Built MSc Quantitative Finance programme",
  "1998-2004", "Currency Trader & Liquidity Manager", "Janus Henderson Investors", "Managed multi-million GBP trading book"
) |>
  knitr::kable(
    col.names = c("Period", "Role", "Organization", "Key Impact"),
    format = "pipe"
  )
```

# Business Impact & Industry Partnerships

## Recent Major Projects (£2M+ Value)

**AI-Driven Regulatory Compliance** (2024-2025) | *£250K UKRI/InnovateUK*
- Partnership with Funds Axis Ltd developing AI solutions for investment management compliance
- Reducing regulatory burden through automated compliance monitoring

**Financial Risk Analytics Platform** (2021-2023) | *£180K InnovateUK KTP*
- Developed advanced tail risk and stress testing analytics with Funds Axis Ltd
- Created commercial-grade VaR modeling and scenario analysis tools

**AI Retail Analytics** (2017-2019) | *£150K InnovateUK KTP*
- Machine learning and predictive analytics for retail business optimization
- Delivered measurable ROI improvements through advanced analytics

## Industry Advisory & Consulting
- **Financial Conduct Authority** Tech Sprint Mentor (2023-present)
- **Citigroup Belfast** - Anomaly detection for trading data (2022-2023)
- **Multiple FinTech partnerships** - RegTech and compliance solutions

# Education & Professional Credentials

```{r}
credentials_industry <- tribble(
  ~Qualification, ~Institution, ~Year, ~Relevance,
  "Ph.D. Finance", "Queen's University Belfast", "2012", "Advanced quantitative finance research",
  "MSc Artificial Intelligence", "Queen's University Belfast", "2025", "Cutting-edge AI applications in finance",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006", "Advanced derivatives and risk management",
  "Chartered Statistician", "Royal Statistical Society", "2019", "Professional statistical expertise"
) |>
  knitr::kable(
    col.names = c("Qualification", "Institution", "Year", "Business Relevance"),
    format = "pipe"
  )
```

# Key Achievements & Recognition

## Awards & Recognition
- **1st Place** - CFA European Quantitative Finance Awards (2018)
- **Advanced Data Science Professional** - Alliance of Data Science Professionals (2024)
- **Teaching Fellow** - Higher Education Authority (2012)

## Research Impact
- **14 peer-reviewed publications** in top-tier finance and technology journals
- **£2M+ research funding** from UKRI, InnovateUK, and industry partners
- **International recognition** - invited speaker at major finance conferences

## Leadership & Innovation
- **Director** - Centre for Financial and Regulatory Technology (2024-present)
- **Co-founder** - Queen's Student Managed Fund (managing real investment portfolio)
- **Programme Director** - MSc Quantitative Finance (2018-2022)

---

# Value Proposition for Industry

## Immediate Impact Areas
1. **AI-Driven Risk Management**: Leverage cutting-edge ML for enhanced risk analytics
2. **Regulatory Technology**: Develop automated compliance and monitoring solutions  
3. **Market Intelligence**: Apply advanced analytics to trading and investment strategies
4. **Strategic Innovation**: Bridge academic research with practical business applications

## Unique Differentiators
- **Proven Industry Track Record**: Real trading floor experience combined with academic rigor
- **Research-to-Practice Pipeline**: Translate cutting-edge research into business solutions
- **Regulatory Expertise**: Deep understanding of financial regulation and compliance
- **Technology Leadership**: At forefront of AI/ML applications in finance

**Available for**: Board positions, strategic consulting, research partnerships, executive roles in FinTech/RegTech

---
*References and detailed project portfolios available upon request*
