---
title: "Prof. Barry Quinn"
subtitle: "Professor of Finance and Financial Technology"
format:
  html:
    template: ../assets/custom-template.html
    css: ../assets/custom-styles.css
    toc: true
    toc-location: right
    toc-title: "Navigation"
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 10pt
    linkcolor: blue
    urlcolor: blue
bibliography: ../data/bqpubs.bib
---

```{r}
#| label: setup 
#| include: FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
library(tidyverse)
library(knitr)
```

# Contact Information

-   **Address**: Ulster University Business School, Belfast Campus, York Street, Belfast, BT1 5AB, N. Ireland
-   **Email**: b.quinn1@ulster.ac.uk
-   **GitHub**: [quinfer](https://github.com/quinfer)

# CV Downloads

Choose the CV version most appropriate for your needs:

::: {.panel-tabset}

## For Academic Positions
**Academic CV** - Focused on research excellence, publications, and academic achievements  
📄 [Download Academic CV (PDF)](cv_academic.pdf) | [View Online](cv_academic.html)

*Ideal for: University positions, tenure applications, academic conferences, research collaborations*

## For Industry Roles  
**Executive Summary** - Business-focused with industry impact and leadership experience  
📄 [Download Industry CV (PDF)](cv_industry.pdf) | [View Online](cv_industry.html)

*Ideal for: Corporate roles, board positions, consulting opportunities, FinTech partnerships*

## For Grant Applications
**Research Portfolio** - Comprehensive research track record and future directions  
📄 [Download Research CV (PDF)](cv_grants.pdf) | [View Online](cv_grants.html)

*Ideal for: Research funding applications, fellowship applications, international collaborations*

## Complete Profile
**Full Resume** - Comprehensive overview of all achievements and experience  
📄 [Download Complete CV (PDF)](index.pdf) | *You are currently viewing the online version*

*Ideal for: General reference, comprehensive overview, personal website*

:::

# Personal Statement

I am Professor of Finance and FinTech at Ulster University Business School and Director of the Centre for Financial and Regulatory Technology. As an applied econometrician and Chartered Statistician, I explore the intersection of finance, technology, and data science, focusing on causal mechanisms beyond cookbook econometrics. My work also engages with computational economics and its finance counterpart, where numerical methods, simulation, and agent-based modelling are applied to problems that defy closed-form solutions.

My research applies statistical, computational, and machine learning techniques to financial markets, causal policy analysis, and cooperative finance. I have been fortunate to receive support from UKRI/EPSRC, Innovate UK, and the Department for the Economy for this work, which has resulted in publications in journals including IEEE Transactions on Evolutionary Computation, International Journal of Forecasting, and Energy Economics.

Before academia, I worked in financial markets as a currency trader and liquidity manager at Janus Henderson Investors, an experience that continues to inform my approach to financial research. I hold a PhD in Finance from Queen's University Belfast and am currently pursuing an MSc in Artificial Intelligence to better understand emerging technologies in finance.

Through my teaching in quantitative finance, econometrics, financial data science, causal AI, and AI in trading, I hope to help students develop both technical competence and ethical approaches to data analysis. By framing my work within computational finance as well as econometrics, I aim to contribute meaningfully to evidence-based decision-making in finance whilst maintaining the rigorous statistical standards expected in our field.

## Education and Professional Qualifications

```{r degrees, results='asis'}
tribble(
  ~Degree, ~Institution, ~Year,
  "B.Sc.(Hons) Accounting and Finance", "Queen's University Belfast", "1995",
  "MSc Quantitative Finance", "RMIT University Melbourne", "2006",
  "Ph.D. Finance", "Queen's University Belfast", "2012",
  "Chartered Statistician", "Royal Statistical Society", "2019",
  "MSc Artificial Intelligence", "Queen's University Belfast", "Graduating 2025"
) |>
  knitr::kable()
```

# Teaching

I have extensive teaching experience at both undergraduate and postgraduate levels, specialising in:

1.  **Quantitative Finance**: My teaching integrates cutting-edge quantitative techniques with practical applications in financial markets.

2.  **Financial Econometrics**: Instructing students in econometrics as an applied statistics discipline, avoiding the statistical cavaliering that plagues the econometric publication playbook.

3.  **Financial Data Science & Causal AI**: End-to-end data pipelines, feature engineering, and time-series ML, with causal inference and causal representation learning for decision support. 
   - Outcomes: reproducible ETL and feature stores, treatment-effect estimation and uplift modeling, model risk management and governance in financial settings.

## Teaching Philosophy

I help students prepare for the future of finance and technology by teaching essential skills like analysing data, making evidence-based decisions under uncertainty, and applying advanced statistical techniques to real-world financial problems. My goal is to equip them with the knowledge and tools they'll need to succeed in a rapidly evolving job market, while also fostering qualities like resilience, humility, and intellectual curiosity.

## Teaching Publication

- Quinn, Barry, Hanna, Alan, Gallagher, Aine (2018). Queen's Student Managed Fund: Investing in the student Experience. Reflections, 27, p16-17


# Research

I have published fourteen peer reviewed internationally acclaimed academic papers (1 ABS4, 11 ABS3s, and 2 ABS2s), three commission research reports, and two statistical software research packages. My research lies at the intersection of finance, economics, technology and data science, with a focus on developing innovative analytical approaches to address complex issues in financial markets, regulation, and sustainability. My work is characterised by the application of advanced statistical, econometric, and machine learning techniques across a diverse range of financial and economic contexts.

## Peer-Reviewed Publications

- [Wang, Hui, Dai, Yongsheng, Spence, Ivor, Rafferty, Karen, Quinn, Barry, Huang, Ji (2025). TDSRL: Time Series Dual Self-Supervised Representation Learning for Anomaly Detection from Different Perspectives. IEEE Internet of Things Journal](https://ieeexplore.ieee.org/document/11048680/authors#authors)

- McKillop, Donal, Liu, Kailong, Quinn, Barry, Peng, Qiao (Forthcoming). Modelling and Predicting Credit Union Failures. *International Journal of Forecasting* (ABS3)

- Liu, Weilong, Zhang, Yong, Liu, Kailong, Quinn, Barry, Yang, Xingyu, Peng, Qiao (2024). Evolutionary Multi-Objective Optimisation for Large-Scale Portfolio Selection With Both Random and Uncertain Returns. *IEEE Transactions on Evolutionary Computation* (ABS4)

- Bouri, E., Quinn, B., Sheenan, L. & Tang, Y. (2024). Investigating extreme linkage topology in the aerospace and defence industry. *International Review of Financial Analysis* (ABS3)

- Quinn, Barry, Gallagher, Ronan, Kuosmanen, Timo (2023). Lurking in the shadows: The impact of CO2 emissions target setting on carbon pricing in the Kyoto agreement period. *Energy Economics* (ABS3)

- Liu, Jiadong, Papailias, Fotis, Quinn, Barry (2021). Direction-of-change forecasting in commodity futures markets. *International Review of Financial Analysis* (ABS3)

- Gallagher, Ronan, Quinn, Barry (2020). Regulatory own goals: The unintended consequences of economic regulation in professional football. *European Sport Management Quarterly* (ABS3)

- McKillop, Donal, French, Declan, Quinn, Barry, Sobiech, Anna L, Wilson, John OS (2020). Cooperative financial institutions: A review of the literature. *International Review of Financial Analysis* (ABS3)

- Quinn, Barry, Hanna, Alan, MacDonald, Fred (2018). Picking up the pennies in front of the bulldozer: The profitability of gilt based trading strategies. *Finance Research Letters* (ABS2)

- McKillop, Donal G, Quinn, Barry (2017). Irish credit unions: Differential regulation based on business model complexity. *The British Accounting Review* (ABS3)

- Ayadi, Rym, Naceur, Sami Ben, Casu, Barbara, Quinn, Barry (2016). Does Basel compliance matter for bank performance? *Journal of Financial Stability* (ABS3)

- Glass, J, McKillop, Donal, Quinn, Barry (2015). Modelling the Performance of Irish Credit Unions, 2002-2010. *Financial Accountability & Management* (ABS3)

- McKillop, Donal G, Quinn, Barry (2015). Web adoption by Irish credit unions: Performance implications. *Annals of Public and Cooperative Economics* (ABS2)

- Glass, J Colin, McKillop, Donal G, Quinn, Barry, Wilson, John (2014). Cooperative bank efficiency in Japan: a parametric distance function analysis. *The European Journal of Finance* (ABS3)

- Quinn, Barry, McKillop, Donal (2009). Internet banking and Irish credit unions. *International Journal of Cooperative Management* (ABS2)

- Quinn, Barry, McKillop, Donal (2009). Cost performance of Irish credit unions. *Journal of Cooperative Studies* (ABS2)

## Software Publications

```{r}
#| label: tbl-software
#| tbl-cap: "Software Publications"

software_table <- tribble(
  ~Package, ~Year, ~Description, ~DOI,
  "Time Series Econometrics (tsfe)", "2022", "R Package for time series analysis", "10.5281/zenodo.6376113",
  "Financial Machine Learning (fml)", "2020", "R Package for financial ML", "10.5281/zenodo.6376114"
)

if (knitr::is_html_output()) {
  software_table %>%
    mutate(
      Package = paste0("[", Package, "](https://doi.org/", DOI, ")"),
      Metrics = case_when(
        str_detect(Package, "tsfe") ~ "[![GitHub stars](https://img.shields.io/github/stars/quinfer/tsfe?style=social)](https://github.com/quinfer/tsfe)",
        str_detect(Package, "fml") ~ "[![GitHub stars](https://img.shields.io/github/stars/quinfer/fml?style=social)](https://github.com/quinfer/fml)"
      )
    ) %>%
    select(Package, Year, Description, Metrics) %>%
    knitr::kable(escape = FALSE)
} else {
  software_table %>%
    mutate(
      Package = paste0(Package, " (DOI: ", DOI, ")")
    ) %>%
    select(Package, Year, Description) %>%
    knitr::kable()
}
```

## Working Paper

- Kearney, Fearghal, Quinn, Barry, Dai, Youngsheng (2025) *Signal Amplification in Market Manipulation Detection*

-   Zhang, Ying (Veronica), Quinn, Barry, Sheenan, Lisa (2025). *Ownership Dynamics, Risk and Regulation in Chinese Banking: New Evidence*. Centre for Responsible Banking & Finance Working Paper Series, WP No. 25-017.

-   Kearney, Fearghal, Quinn, Barry, Pramanick, Abhishek (Under Review at Journal of Business Venturing) *Innovation Signals and Strategic Exits: How Technological Readiness Shapes RegTech Venture Outcomes*

- Quinn, Barry, Gallagher, Ronan *Great Expectations: Managerial Turnover and Market Expectation in Association Football*


## Research Reports

-   Bryon Graham, Barry Quinn (2019). *Price comparison and web analytics*. (Commissioned by InvestNI)
-   Bryon Graham, Barry Quinn (2017). *Machine learning and predictive analytics in the mall retail business*. (Commissioned by InvestNI)
-   Fearghal Kearney, Barry Quinn (2020). The theoretical foundations of value at risk modelling. (Commissioned by InvestNI and Funds Axis Ltd)
-   French,Declan, McKillop, Donal, Quinn, Barry (2018). *Landscape review of Northern Ireland Credit Union Sector.*(Commissioned by the NI Department of Communities and Rural Affairs)
- McKillop, Donal, Quinn, Barry (2012 )Report of the Irish Commission on Credit Unions (4* REF2014 Impact Case): An Efficiency Study

## Research Impact and Knowledge Transfer

1.  *Understanding and Enhancing regulatory compliance in the Investment Managent industry using AI* with Funds Axis Ltd and Momentum 1.0 (Funded by UKFin+ and UKRI Nov 2024 - Nov 2025)
2.  *Anomaly detection of large heterogeneous trading transaction and commmunication data* with Citigroup Belfast and Momentum 1.0 (2022-2023)
3.  *AI and Advanced Retail Analytics* with Pearlai Ltd and Dr Byron Graham (KTP 2017-2019)
4.  *E.S.G Fair Value Analytics* with Research Associates, Dr. Lisa Sheenan, Dr. Byron Graham. (KTP 2021)
5.  *Tail Risk Analytics, Stress Testing and Scenario Analytics* with Funds Axis Ltd, Dr Fearghal Kearney, and Dr Colm Kelly. (KTP 2021 -2023)
6.  *Financial Conduct Authority* Tech Sprint Mentor (2023-present)
7.  *AI and the Future of Financial Regulation* with Fearghal Kearney and Abhishek Praminick published in The Economic Observatory (2023)

## Completed PhD Supervisions

1.  Dr. Jiadong liu (graduated 2018): *Momentum in Empirical Asset Pricing* **(Principal Supervisor)**
2.  Dr. Ashleigh Neil (graduated 2019): *Law and Financial Stability* **(Principal Supervisor)**
3.  Dr. Colm Kelly (graduated 2021): *Machine learning in Empirical Asset Pricing* **(Principal Supervisor)**
4.  Veronica Zhang (graduating 2024): *Capital policy and State sponsorship in Chinese banking* **(Principal Supervisor)**
5.  Dr. Kevin Johnson (graduated 2019) *A Case Study in Industrial Consolidation: The Irish Credit Union Sector* **(Co Supervised)**
6.  Dr. Qiao (Olivia) Peng (graduated 2021) *US Credit Union Mergers: Causes and Consequences* **(Co Supervised)**

## Current PhD Supervisions

1.  Fei Li *Sustainable Investing, Report and Gender Diversity* (fully-funded by QBS scholarship) **(Principal Supervisor)**
2.  Abhishek Pramnanick *AI and the future of financial regulation* (fully-funded by QBS scholarship) **(Principal Supervisor)**
3.  Dongwei (Jeffrey) Li *The Economics of AI patents and productivity in finance* (fully-funded by DofE scholarship) **(Principal Supervisor)**
4.  Yuqi Ding *The Economics of Generative AI in Finance* (fully-funded by QBS scholarship) **(Principal Supervisor)**
5.  Yongsheng Dai "Market Manipulation and Financial Anomaly Detection" (fully-funded by PWC scholarship) **(Co Supervisor)**
6.  Brandon Cochrane "Economic Cost of Cultural Displays in Northern Ireland" (fully-funded by DofE CAST scholarship) **(Co Supervisor)**

## Visiting PhD Supervision

1.  Weilong Liu 2022 (From Lingnan University College of Yat-sen University (SYSU) in Guangzhou, China ) "Market Manipulation and AI Anomaly Detection"
2.  James Hannon 2023 (From UCD) "Leveraging Computer Vision to under and enhance Insurance Pricing"
3.  Birem Abderraouf 2025 (From UPEC Paris) "Generative AI in Capital Markets"

## Conference presentations

-   Invited Talk, *Rethinking Research Impact: Combining Effect Sizes and Economic Significance in Finance*,Royal Statistics Society Annual Conference, Brighton UK, Sep 2024
-   Keynote Speaker, *Estimating Systemic Risk*, Irish Finance Association, Maynooth University, Ireland, April 2023
-   Invited Talk, *Teaching data science in the age of FinTech*, Royal Statistics Society Annual Conference, Aberdeen UK, Sep 2022
-   Invited Talk, *Carbon pricing and machine learning*, Multidisciplinary Workshop on Fintech, Islamic Finance and Sustainability (online), Hamad Bin Khalifa University, Qatar, Nov 2022
-   Invited Speaker, *Understanding fintech and financial stability*, International Workshop on Financial System Architecture and Stability, Bayes Business School, London, Sept 2018
-   Invited Speaker, *Systemic Risk and Basel Compliance*, British Accounting and Finance Association Annual Conference, London, April 2018
-   Invited Speaker, *Differential regulation of Irish credit unions: Does one size fit all?*, 2nd Conference on Contemporary Issues in Banking, Centre for Responsible Banking and Finance, St Andrews, Dec 2017
-   Invited Speaker, *Business model diversity, efficiency and productivity of cooperatives*, European workshop in efficiency and productivity analysis, Aalto University, Finland, Jun 2017
-   Invited Panelist, *Statistical inference and credibility in finance*, Emerging Scholars in Banking and Finance, Bayes Business School, London, Dec 2016
-   Invited Speaker, *Capital regulation compliance and the performance of European banks*, International Workshop on Financial System Architecture and Stability, HEC Montreal, Aug 2016
-   Participant, *Bloomberg Annual Educational Symposium*, Bloomberg London HQ, Sep 2015

```{r}
#| label: tbl-research-grants
#| tbl-cap: "Research Grant Income"
if (knitr::is_latex_output()) {
  read.csv("../data/Grant_income.csv") %>%
    mutate(Amount = scales::comma(Amount, prefix = "£")) %>%
    knitr::kable(
      col.names = c("Source", "Role", "Start", "End", "People", "Title", "Full Economic Costs (£)", "Type"),
      align = c("l", "l", "l", "l", "l", "l", "r", "l"),
      format = "latex",
      booktabs = TRUE,
      escape = FALSE,
      longtable = TRUE
    ) %>%
    kableExtra::kable_styling(
      font_size = 9,
      latex_options = c("repeat_header")
    ) %>%
    kableExtra::column_spec(1, width = "1.5cm") %>%  # Source
    kableExtra::column_spec(2, width = "1.5cm") %>%  # Role
    kableExtra::column_spec(3, width = "1.2cm") %>%  # Start
    kableExtra::column_spec(4, width = "1.2cm") %>%  # End
    kableExtra::column_spec(5, width = "2cm") %>%    # People
    kableExtra::column_spec(6, width = "4cm") %>%    # Title
    kableExtra::column_spec(7, width = "2cm") %>%    # Amount
    kableExtra::column_spec(8, width = "1.5cm")      # Type
} else {
  read.csv("../data/Grant_income.csv") %>%
    mutate(Amount = scales::comma(Amount, prefix = "£")) %>%
    knitr::kable(
      col.names = c("Source", "Role", "Start", "End", "People", "Title", "Full Economic Costs (£)", "Type"),
      align = c("l", "l", "l", "l", "l", "l", "r", "l"),
      format = "pipe",
      caption = "Research Grant Income",
      escape = FALSE
    ) %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left"
    )
}
```

# Citizenship

## Academic Service

-   **Co-director**, Finance and AI Research lab, Queen's University Belfast, 2022--2024.
-   **Lead Developer**, Queen's Business School Remote Analytics Lab Platform (Q-RaP), 2021--2024.
-   **QUB Academic Lead**, Steering group for Northern Ireland global centre in secure connected intelligence for regulatory technology in finance, 2022--2024.
-   **Programme Director**, MSc Quantitative Finance, Queen's University Belfast, 2018--2022.
-   **Programme Director**, MSc Computational Finance & Trading, Queen's University Belfast, 2014--2018.
-   **Co-Founder**, Queen's Student Managed Fund, 2012--2024.
-   **Organiser** of Symposium on AI and the Future of Financial Regulation, Queen's University Belfast, November 2023
-   **Organiser** of Symposium on Safety and Assurance in Finance. Queen's Business School, November 2024
-   **Chair**, Panel debate on "Digital Regulation: Shaping digital markets and safeguarding consumer rights in Northern Ireland", Northern Ireland Competition Forum, May 2024

## Professional Accreditations

-   Chartered Statistician, Royal Statistical Society (2019)
-   Advanced Data Science Professional, [Alliance of Data Science Professional](https://alliancefordatascienceprofessionals.com/) (2024)

# Awards and Recognition

-   Teaching Fellow of Higher Education Authority (2012)
-   QUB Teaching Award (2016)
-   1^st^ place in CFA European Quantitative Finance Awards (2018)
-   EEECS Scholarship for AI MSc (2023)
-   Associate Research Fellow, QUB Momentum One Zero (formerly Global Innovation Institute) (2021--2024)

# Professional Experience

```{r}
employment_table <- tribble(
  ~Start, ~End, ~Position, ~Institution,
  1995, "1998", "Financial Adviser", "City Financial Partners, London",
  1998, "2004", "Currency Trader and Liquidity Manager", "Janus Henderson Investors, London",
  2005, "2005", "Teaching Fellow and Quantitative Researcher", "Department of Finance, RMIT University Melbourne",
  2009, "2009", "Teaching Fellow", "Business School, Ulster University",
  2010, "2020", "Lecturer", "Department of Finance, Queen's Management School",
  2020, "2024", "Senior Lecturer in Finance, Technology and Data Science", "Queen's Management School",
  2024, "present", "Professor of Finance and FinTech", "Ulster University Business School",
  2024, "present", "Director, Centre for Financial and Regulatory Technology", "Ulster University Business School"
) |>
  arrange(desc(Start)) |>
  mutate(
    Start = as.character(Start),
    Years = if_else(End == "present", 
                   paste0("**", Start, " - Present**"), 
                   paste0("**", Start, " - ", End, "**"))
  ) |>
  select(Years, Position, Institution)

if (knitr::is_latex_output()) {
  employment_table |>
    knitr::kable(
      col.names = c("Period", "Position", "Institution"),
      format = "latex",
      caption = "Professional Experience",
      booktabs = TRUE,
      escape = FALSE
    ) |>
    kableExtra::kable_styling(
      latex_options = "scale_down",
      font_size = 9
    )
} else {
  employment_table |>
    knitr::kable(
      col.names = c("Period", "Position", "Institution"),
      format = "pipe",
      caption = "Professional Experience",
      escape = FALSE
    ) |>
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left"
    )
}
```

# Skills and Expertise

-   Advanced statistical modeling and machine learning
-   Ethical econometrics
-   AI applications in finance and trading
-   Economics of AI and machine learning
-   Financial risk management and systemic risk analysis
-   Data science and big data analytics in finance
-   Interdisciplinary research leadership
